---
title: "Introduction to the raustats package"
author: "David Mitchell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{01_-_Using_the_raustats_package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


# Introduction

The [Australian Bureau of Statistics (ABS)](http://www.abs.gov.au/) is
Australia’s national statistical agency, providing trusted official
statistics on a wide range of economic, social, population and
environmental matters of importance to Australia. Key ABS statistical
collections include:
 
  * Australian National Accounts
  * International Trade
  * Consumer Price Index (CPI)
  * Labour Force
  * Population trends


The [Reserve Bank of Australia (RBA)](http://www.rba.gov.au/) is
Australia's central bank. In addition to its legislative
responsibilities, it collects and publishes statistics on money,
credit, the Australian banking systems and other relevant economic
metrics. Key RBA statistics include:

  * Banking system assets and liabilities
  * Money and credit statistics
  * Household and business finances
  * Interest rates
  * Exchange rates
  * Inflation and inflation expectations.

The ABS and RBA currently make their statistics primarily available
through Excel and/or CSV spreadsheets. This is not machine readable,

... The functions
<!-- UP TO HERE --> [TO BE COMPLETED]



The ABS also has a Beta API access portal:
[ABS.Stat](http://stat.data.abs.gov.au/), which provides access to a
truncated range of ABS data and statistics. This package also provides
some functions to search and download data from the ABS.Stat
API. Please note that as the API is currently in development, changes
in the API backend may break the functions provided in this package.

The `raustats` package allows researchers to quickly search and
download ABS and RBA data of interest in a programmatic and
reproducible fashion. This facilitates seamless integration into
projects and enables analysis to be more seamlessly re-run using the
latest available data.


### Main features of the `raustats` package: 

  * Access to all time series data in the ABS statistical catalogue
    and the RBA website.
  * Access to the ABS Beta API: [ABS.Stat](http://stat.data.abs.gov.au) 
  * Ability to return `POSIXct` dates for easy integration into
    plotting and time-series analysis techniques 
  * Returns data in long (or tidy) format for direct
    integration with packages like `ggplot2`, `tidyr` and `dplyr`
  * Support for `grep` style searching for data descriptions and names.



# Quick start guide

## Downloading ABS Catalogue data


## Downloading ABS data via [ABS.Stat](https://stat.data.abs.gov.au/)



## Downloading RBA data




# ABS data access functions

##  ABS Catalogue data functions



##  ABS API data access functions - Beta 



# RBA data access functions



<!-- UP TO HERE -->

Unless you know the country and indicator codes that you want to
download the first step would be searching for the data you are
interested in. `abs_search()` provides `grep` style searching of all
available indicators from the ABS API and returns the indicator
information that matches your query.

To access what countries or regions are available you can use the
`countries` data frame from either `abs_cachelist` or the saved return
from `abs_cache()`. This data frame contains relevant information
regarding each country or region. More information on how to use this
for downloading data is covered later.


## Finding available data with `abs_cachelist`

For performance and ease of use, a cached version of useful
information is provided with the `wbstats` R-package. This data is
called `wb_cachelist` and provides a snapshot of available countries,
indicators, and other relevant information. `wb_cachelist` is by
default the the source from which `wbsearch()` and `wb()` uses to find
matching information. The structure of `wb_cachelist` is as follows
```{r eval=FALSE}
library(raustats)

str(wb_cachelist, max.level = 1)
```


## Accessing updated available data with `wbcache()`

For the most recent information on available data from the World Bank
API `wbcache()` downloads an updated version of the information stored
in `wb_cachelist`. `wb_cachelist` is simply a saved return of
`wbcache(lang = "en")`. To use this updated information in
`wbsearch()` or `wb()`, set the `cache` parameter to the saved `list`
returned from `wbcache()`. It is always a good idea to use this
updated information to insure that you have access to the latest
available information, such as newly added indicators or data sources.

```{r eval=FALSE}
library(raustats)

# default language is english
new_cache <- wbcache()
```


## Search available data with `wbsearch()`

`wbsearch()` searches through the `indicators` data frame to find
indicators that match a search pattern. An example of the structure of
this data frame is below

```{r, echo=FALSE, results='asis'}
knitr::kable(head(wbstats::wb_cachelist$indicators[4310:4311, ]))
```

By default the search is done over the `indicator` and `indicatorDesc`
fields and returns the columns `indicatorID` and `indicator` of the
matching rows. The `indicatorID` values are inputs into `wb()`, the
function for downloading the data. To return all columns for the
`indicators` data frame, you can set `extra = TRUE`.

```{r eval=FALSE}
library(wbstats)

unemploy_vars <- wbsearch(pattern = "unemployment")
head(unemploy_vars)

```

Other fields can be searched by simply changing the `fields` parameter. For example
```{r eval=FALSE}
library(raustats)

blmbrg_vars <- wbsearch(pattern = "Bloomberg", fields = "sourceOrg")
head(blmbrg_vars)

```

Regular expressions are also supported.
```{r eval=FALSE}
library(raustats)

# 'poverty' OR 'unemployment' OR 'employment'
povemply_vars <- wbsearch(pattern = "poverty|unemployment|employment")

head(povemply_vars)

```

The default cached data in `wb_cachelist` is in English. To search
indicators in a different language, you can download an updated copy
of `wb_cachelist` using `wbcache()`, with the `lang` parameter set to
the language of interest and then set this as the `cache` parameter in
`wbsearch()`. Other languages are supported in so far as they are
supported by the original data sources. Some sources provide full
support for other languages, while some have very limited support. If
the data source does not have a translation for a certain field or
indicator then the result is `NA`, this may result in a varying number
matches depending upon the language you select.

```{r eval=FALSE}
library(raustats)

# download wbcache in spanish
wb_cachelist_es <- wbcache(lang = "es")

gini_vars <- wbsearch(pattern = "Coeficiente de Gini", cache = wb_cachelist_es)

head(gini_vars)

```

## Downloading data with `wb()`

Once you have found the set of indicators that you would like to explore further, the next step is downloading the data with `wb()`. The following examples are meant to highlight the different ways in which `wb()` can be used and demonstrate the major optional parameters.

The default value for the `country` parameter is a special value of `all` which as you might expect, returns data on the selected `indicator` for every available country or region.
```{r eval=FALSE}
library(raustats)

# Population, total
pop_data <- wb(indicator = "SP.POP.TOTL", startdate = 2000, enddate = 2002)

head(pop_data)
```

If you are interested in only some subset of countries or regions you can pass along the specific codes to the `country` parameter. The country and region codes that can be passed to the `country` parameter correspond to the coded values from the `iso2c`, `iso3c`, `regionID`, `adminID`, and `incomeID` from the `countries` data frame in `wb_cachelist` or the return of `wbcache()`. Any values from the above columns can mixed together and passed to the same call
```{r eval=FALSE}
library(raustats)

# Population, total
# country values: iso3c, iso2c, regionID, adminID, incomeID
pop_data <- wb(country = c("ABW","AF", "SSF", "ECA", "NOC"),
               indicator = "SP.POP.TOTL", startdate = 2012, enddate = 2012)

head(pop_data)
```

Queries with multiple indicators return the data in a long data format by default
```{r eval=FALSE}
library(raustats)

pop_gdp_long <- wb(country = c("US", "NO"), indicator = c("SP.POP.TOTL", "NY.GDP.MKTP.CD"),
                   startdate = 1971, enddate = 1971)

head(pop_gdp_long)
```

or a wide format if parameter `return_wide = TRUE`. Note that to necessitate a this transformation the `indicator` column is dropped.
```{r eval=FALSE}
library(raustats)

pop_gdp_wide <- wb(country = c("US", "NO"), indicator = c("SP.POP.TOTL", "NY.GDP.MKTP.CD"),
                   startdate = 1971, enddate = 1971, return_wide = TRUE)

head(pop_gdp_wide)
```

### Using `mrv`
If you do not know the latest date an indicator you are interested in is available for you country you can use the `mrv` instead of `startdate` and `enddate`. `mrv` stands for most recent value and takes a `integer` corresponding to the number of most recent values you wish to return
```{r eval=FALSE}
library(raustats)

eg_data <- wb(country = c("IN"), indicator = 'EG.ELC.ACCS.ZS', mrv = 1)

eg_data
```

You can increase this value and it will return no more than the `mrv` value. However, if `mrv` is greater than the number of available data it will return less
```{r eval=FALSE}
library(raustats)

eg_data <- wb(country = c("IN"), indicator = 'EG.ELC.ACCS.ZS', mrv = 10)

eg_data
```

### Using `gapfill = TRUE`

An additional parameter that can be used along with `mrv` is
`gapfill`. `gapfill` allows you to "fill-in" the values between actual
observations. The "filled-in" value for an otherwise missing date is
the last observed value carried forward.The only difference in the
data call below from the one directly above is `gapfill = TRUE` (the
default is `FALSE`). Note the very important difference

```{r eval=FALSE}
library(raustats)

eg_data <- wb(country = c("IN"), indicator = 'EG.ELC.ACCS.ZS', mrv = 10, gapfill = TRUE)

eg_data
```

Because `gapfill` returns data that does reflect actual observed
values, use this option with care.

### Using `POSIXct = TRUE`

The default format for the `date` column is not conducive to sorting
or plotting, especially when downloading sub annual data, such as
monthly or quarterly data. To address this, if `TRUE`, the `POSIXct`
parameter adds the additional columns `date_ct` and
`granularity`. `date_ct` converts the default date into a
`POSIXct`. `granularity` denotes the time resolution that the date
represents. This option requires the use of the package `lubridate (>=
1.5.0)`. If `POSIXct = TRUE` and `lubridate (>= 1.5.0)` is not
available, a `warning` is produced and the option is ignored

```{r eval=FALSE}
library(raustats)

oil_data <- wb(indicator = "CRUDE_BRENT", mrv = 10, freq = "M", POSIXct = TRUE)

head(oil_data)
```

The `POSIXct = TRUE` option makes plotting and sorting dates much easier.
```{r vig-ex-posixct-1, fig.height = 4, fig.width = 7.5, fig.path = "figures/"}
library(raustats)
library(ggplot2)

oil_data <- wb(indicator = c("CRUDE_DUBAI", "CRUDE_BRENT", "CRUDE_WTI", "CRUDE_PETRO"),
               startdate = "2012M01", enddate = "2014M12", freq = "M", POSIXct = TRUE)

ggplot(oil_data, aes(x = date_ct, y = value, colour = indicator)) + geom_line(size = 1) +
  labs(title = "Crude Oil Price Comparisons", x = "Date", y = "US Dollars")
```


The `POSIXct = TRUE` option also makes plotting time series with
different time coverage seamless.

```{r vig-ex-posixct-2, fig.height = 4, fig.width = 7.5, fig.path = "figures/"}
library(raustats)
library(ggplot2)

# querying seperate for differing time coverage example
gold_data <- wb(indicator = "GOLD", mrv = 120, freq = "M", POSIXct = TRUE)
plat_data <- wb(indicator = "PLATINUM", mrv = 60, freq = "M", POSIXct = TRUE)

metal_data <- rbind(gold_data, plat_data)

ggplot(metal_data, aes(x = date_ct, y = value, colour = indicator)) + geom_line(size = 1) +
  labs(title = "Precious Metal Prices", x = "Date", y = "US Dollars")
```


# Some Sharp Corners

There are a few behaviors of the World Bank API that being aware of
could help explain some potentially unexpected results. These results
are known but no special actions are taken to mitigate them as they
are the result of the API itself and artifically limiting the inputs
or results could potentially causes problems or create unnecessary
rescrictions in the future.


## Most Recent Values

If you use the `mrv` parameter in `wb()` with mutliple countries or
regions, it searches for the most recent dates for which any country
or region in your selection has data and then returns the data for
those dates. In other words the `mrv` value is not determined on a
country by country basis, rather it is determined across the entire
selection.

```{r eval=FALSE}
library(raustats)

per_data_1 <- wb(country = "all", indicator = 'per_lm_ac.cov_pop_tot', mrv = 1)
per_data_1

per_data_2 <- wb(country = "all", indicator = 'per_lm_ac.cov_pop_tot', mrv = 2)
per_data_2

```

## Searching in other languages
Not all data sources support all languages. If an indicator does not have a translation for a particular language, the non-supported fields will return as `NA`. This could potentially result in a differing number of matching indicators from `wbsearch()`
```{r eval=FALSE}

library(raustats)

# english
cache_en <- wbcache()
sum(is.na(cache_en$indicators$indicator))

```





