---
title: "Introduction to the raustats package"
author: "David Mitchell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{01_-_Using_the_raustats_package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


# Introduction

The [Australian Bureau of Statistics (ABS)](http://www.abs.gov.au/) is
Australia’s national statistical agency, providing trusted official
statistics on a wide range of economic, social, population and
environmental matters of importance to Australia. Key ABS statistical
collections include:
 
  * Australian National Accounts
  * International Trade
  * Consumer Price Index (CPI)
  * Labour Force
  * Population trends


The [Reserve Bank of Australia (RBA)](http://www.rba.gov.au/) is
Australia's central bank. In addition to its legislative
responsibilities, it collects and publishes statistics on money,
credit, the Australian banking systems and other relevant economic
metrics. Key RBA statistics include:

  * Banking system assets and liabilities
  * Money and credit statistics
  * Household and business finances
  * Interest rates
  * Exchange rates
  * Inflation and inflation expectations.

The ABS and RBA currently make most, or all, of their statistics primarily
available through Excel and/or CSV-format spreadsheets. The former typically
require some additional re-formatting to produce well-formed (tidy) data. The
functions in this package automate translation the downloading and formatting of
ABS and RBA data into long-format (tidy) tables.

The ABS also has a Beta API access portal:
[ABS.Stat](http://stat.data.abs.gov.au/), which provides access to a subset of
all ABS statistics. This package also provides a reasonably comprehensive range
of functions to search and download data from the [ABS.Stat](http://stat.data.abs.gov.au/) API. Please note that
as the API is currently in development, changes in the API backend may break the
functions provided in this package.

The `raustats` package allows researchers to quickly search and download ABS and
RBA data of interest in a programmatic and reproducible fashion. This
facilitates seamless integration into projects and enables analysis to be more
seamlessly re-run using the latest available data.


### Main features of the `raustats` package: 

  * Access to all time series and cross-section statistics in the ABS
    statistical catalogue and on the RBA website
  * Access to statistics available through the ABS Beta API:
    [ABS.Stat](http://stat.data.abs.gov.au)
  * Ability to return `POSIXct` dates for easy integration into
    plotting and time-series analysis techniques 
  * Returns data in long (or tidy) format for direct
    integration with packages like `ggplot2`, `tidyr` and `dplyr`
  * Support for regular expression (`grep`) style searching for data
    descriptions and names.


# Quick-start guide

This section provides a quick-start guide on how to download ABS
and RBA statistics. First, load the library:

```{r eval=FALSE}
library(raustats)
```

## Downloading ABS Catalogue Statistics

The `abs_cat_stats()` function is the workhorse for downloading ABS statistics
by ABS Catalogue Number. Downloading the latest  Consumer Price Index (CPI) data
series (ABS Catalogue no. 6401.0) can be simply done with the following call:

```{r eval=FALSE}
cpi_all <- abs_cat_stats("6401.0")
```

To download the latest statistics reported in Table 1 (ABS groups Tables 1 and
2), simply provide a regular expression to the `tables` argument:

```{r eval=FALSE}
cpi <- abs_cat_stats("6401.0", tables="Table.+1")
```

Statistics from previous releases can be accessed via the `releases` argument. A
more detailed explanation of the `abs_cat_stats()` function is provided below.


## Downloading ABS data via [ABS.Stat](https://stat.data.abs.gov.au/)

The `abs_stats()` function is used to download ABS statistics via the ABS.Stat
API. As at the time of writing, the [ABS.Stat](http://stat.data.abs.gov.au/) API had a limit on query string
length and number of records returned.  In order to download the latest CPI data
series for All Group changes in prices across Australia and each of the eight
capital cities, simply call:

```{r eval=FALSE}
cpi_api <- abs_stats("CPI", filter=list(MEASURE=1, REGION=c(1:8,50), INDEX=10001, TSEST=10, FREQUENCY="Q"))
```

The package includes additional functions to search for datasets and data series
available on ABS.Stat. These are documented further below.


## Downloading RBA data

The `rba_stats()` function is the workhorse for downloading statistics available
on the RBA website.  Downloading the latest statistics covering the RBA's assets
and liabilities (RBA Statistical Table A1), can be simply done with the
following call:

```{r eval=FALSE}
rba_bs <- rba_stats("A1")
```


# ABS statistics access functions
##  ABS Catalogue statistics functions

The ABS Catalogue statistics functions are split into core functions:
  * `abs_cat_stats`
  * `abs_cat_tables`
  
and helper functions:
  * `abs_read_tss`
  * `abs_download_data`
  * `abs_unzip_files`

Helper functions are called by the core functions and should not need to be
accessed directly by users.


### Finding available ABS Catalogue statistics

The ABS does not provide a single consolidated list of former and current
statistical collections available through the ABS Catalogue. A text search
facility is available on the ABS website:
[www.abs.gov.au](http://www.abs.gov.au), but this package does not provide any
functionality to access this. Instead, the `abs_tablecache` data set contained
in this package lists the more common ABS Catalogue statistics.


### Accessing ABS Catalogue statistics with `abs_cat_stats()`

As previously outlined in the Quick Start section, the
`abs_cat_stats()` function provides easy access to ABS statistics by
Catalogue Number.  The following examples demonstrate typical uses of
the function and the various function arguments.

The simplest use of the `abs_cat_stats()` function is to download all
tables available in the specified catalogue number series. The Quick
Start illustrated how to download all CPI tables. The following
example downloads the latest quarterly national accounts statistics
(ABS Catalogue no. 5206.0):

```{r eval=FALSE}
# Latest quarterly national accounts
ana_q <- abs_cat_stats(series = "5206.0")
head(ana_q)
```

The function returns a long-format (tidy) table with the following
columns:
  * series_id -- ABS series identifier
  * date -- Date-format date
  * value -- series value
  * data_item_description -- data item name and description
  * series_type -- series type, one of Original, Trend, Seasonally Adjusted
  * series_start -- series start date
  * series_end -- series end data
  * no_obs -- number of series observations
  * unit -- unit type (e.g. Percent, $ Millions, Index Numbers, Proportion)
  * data_type -- data type (e.g. Derived)
  * freq (frequency) -- Frequency (e.g. Annual, Quarterly, Monthly)
  * collection_month -- collection month (integer)
  * catalogue_no -- catalogue number (e.g. 5206.0, 6401.0)
  * publication_title -- ABS publication title
  * table_no -- ABS publication table number (integer)
  * table_title -- ABS publication table title.


The `tables` argument allows users to select the set of catalogue
tables to be downloaded by specifying a regular expression to pattern
match the ABS table names, as specified on the ABS web page for the
specified Catalogue number. By default
`tables = "All"`, which automatically downloads all available tables from the
specified catalogue number. For example, the following code downloads
Tables 1 and 2 from Catalogue no. 5206.0.

```{r eval=FALSE}
# Latest quarterly national accounts
ana_q <- abs_cat_stats(series = "5206.0", tables=c("^Table 1\\.", "^Table 2\\."))
head(ana_q)
```

<!-- The same result may be achieved by specifying the website table -->
<!-- names, e.g.: -->

```{r eval=FALSE, echo=FALSE}
# Latest quarterly national accounts
ana_q <- abs_cat_stats(series = "5206.0",
                       tables=c(".*Key National Accounts Aggregates",
                                ".*Expenditure on Gross Domestic Product (GDP), Chain volume measures"))
head(ana_q)
```

The function `releases` argument enables users to download data from
the specified release. By default, the function downloads the latest
available data (i.e. `releases="Latest"`). The format is a date object
or character string specifying the month and year of release. 
For example, the following sample code downloads Table 1 from the
December 2017 release of the quarterly national accounts. The
`releases` argument accepts multiple 

```{r eval=FALSE}
## December 2017 quarter national accounts - Table 1
ana_2017Q4 <- abs_cat_stats("5206.0", tables="Table 1", releases="Dec 2017");
## or
ana_2017Q4 <- abs_cat_stats("5206.0", tables="Table 1", releases=as.Date("2017-12-01"));
head(ana_2017Q4)
```

The `abs_cat_stats` can download both time series (`types="tss"`) and
cross-section (`types="css"`) data series.  ABS time series
spreadsheets generally have a standard format, with a single column
for each series, with several headers rows containing series metadata,
including a single row with the unique series identifier.  ABS
cross-section spreadsheet formats vary, depending on the number of
dimensions (categories) available in the data set. These tables
typically have multiple uniquely-identifying header rows. Importing
these tables is handled by specifying `types="css"` and supplying the
header start row (e.g. `header_row=4`, if the series header
information starts in row 4) and the number of header rows
(e.g. `header_rows=2`, if there are two header rows). The following
example shows the use of the function to import cross-section data on
regional household incomes (ABS Catalogue no. xxxx.0).

```{r eval=FALSE}
## -- TO BE ADDED --
head(reg_inc)
```


### Finding available ABS Catalogue tables with `abs_cat_tables`

The `abs_cat_tables` function returns a list of all tables for one or more
specified ABS Catalogue numbers.  It can be used to identify the relevant table
to download. For example, the following code returns all tables available for
the latest quarterly national accounts (5206.0) and CPI (6401.0).


```{r eval=FALSE}
# Latest quarterly national accounts
ana_tables <- abs_cat_tables(cat_no="5206.0")
head(ana_tables)

cpi_tables <- abs_cat_tables(cat_no="6401.0")
head(cpi_tables)
```

The `abs_cat_tables` also has three additional arguments: `releases`,
`types` and `include_urls`.  The `releases` argument returns the list
of downloadable tables from the specified release. Setting
`releases="Latest"` (the default) returns the latest available tables.
Earlier release tables can be returned by specifying the month and
year of release, e.g.`releases = "Jun 2017"`.  The `types` argument enables users to specify which file types to
include. Options are 'tss' -- ABS Time Series Spreadsheets, 'css' -
ABS Data Cubes and 'pub' -- ABS Publications. The default is`types = c('tss', 'css')`.  The `include_urls` argument specifies whether or not to include the
URLs of available data tables in the returned results. The default is
`include_urls=FALSE`.

The following example, returns all quarterly national accounts tables in the
September and December 2017 quarter releases, and includes the file paths.

```{r eval=FALSE}
# Specified quarterly national accounts tables
ana_tables <- abs_cat_tables(cat_no="5206.0", releases=c("Sep 2017", "Dec 2017"), include_urls=TRUE)
head(ana_tables)
```

And the following example illustrates use of the `abs_cat_tables` to get all
available downloadable Data Cubes for a non-time series collection, the
Australian Statistical Geography Standard (ASGS) main structure classification
and digital boundaries (Catalogue no. 1270.0.55.001).

```{r eval=FALSE}
# Specified ASGS files
asgs_files <- abs_cat_tables(cat_no="1270.0.55.001", types="css", include_urls=TRUE)
head(asgs_files)
```


### Other ABS Catalogue helper functions

There are several ABS Catalogue helper functions that are called by
`abs_cat_stats` and `abs_cat_tables` ---that download and parse the
ABS Catalogue table files. These additional functions should not
ordinarily need to be called directly, but there may be some
situations in which users wish to access these functions. The main
ones are:

  * `abs_cat_download`
  * `abs_cat_unzip`
  * `abs_read_tss`
  * `abs_read_css`

The following examples illustrate the use of these functions.

The `abs_cat_download` function downloads and saves ABS Catalogue
tables from a supplied URL. It is called inside the `abs_cat_stats`
and can be used to directly download one or more ABS Catalogue table
files. It is most usefully used in conjunction with the
`abs_cat_tables` function, as follows:

```{r eval=FALSE}
# Read specified ABS Excel Time Seriefiles
tables <- abs_cat_tables("5206.0", releases="Latest", include_urls=TRUE);
downloaded_tables <- abs_cat_download(tables$path_2[1], exdir=tempdir())

print(downloaded_tables)
```

The `abs_cat_unzip` function extracts Excel files from compressed ABS zip
archives, it uses the `utils::unzip` function, but wraps around some standard
file locations. There are two arguments: `files` and `exdir` which have a
similar meaning to the `utils::unzip` equivalent arguments. By default
`exdir = tempdir()`.

```{r eval=FALSE}
# Extracted ABS Excel Time Series Spreadsheet
extracted_files <- abs_cat_unzip(downloaded_tables)

print(extracted_files)
```

The `abs_read_tss`, and companion `abs_read_tss_`, functions are among the more
important functions in the package. It extracts data from standard-formatted ABS
Catalogue time series spreadsheets and returns it as a long-format (tidy) data
frame.  The simple example below shows use of the function to read Catalogue
5206.0, Table 1.

```{r eval=FALSE}
# Read specified ABS Excel Time Series Spreadsheets
x <- abs_read_tss(extracted_files);
head(x)
```


##  ABS.Stat statistics access functions - Beta 

[ABS.Stat](http://stat.data.abs.gov.au/) statistics access functions
available in `raustats` include functions to list and search for
available data sets and statistics, and functions to download selected
statistics.


### Finding available data with `abs_datasets`

The `abs_datasets` function returns a list of all datasets available
through ABS.Stat. The function has two arguments: `lang` (default is
English: `lang="en"`) and `include_notes` (default:
`include_notes=FALSE`). The following example shows the results with
notes included.

```{r eval=FALSE}
datasets <- abs_datasets(include_notes=TRUE);

head(datasets)
```

### Cached list of available datasets `abs_cachelist`

For performance, a cached list of datasets available through the [ABS.Stat](http://stat.data.abs.gov.au/) API is
provided in the `abs_cachelist` data set included in `raustats`.
`abs_cachelist` is the default source used in `abs_search()` and `abs_stats()`
to find matching dataset information. The structure of `abs_cachelist` is shown
next.
```{r eval=FALSE}
str(abs_cachelist, max.level = 1)
```

The default cached data in `abs_cachelist` is in English. To search indicators
in a different language, you can download an updated copy of `abs_cachelist`
using `abs_cache()` (see next).


### Accessing updated available data with `abs_cache()`

Current information about datasets available from the [ABS.Stat](http://stat.data.abs.gov.au/) API can be
obtained by calling `abs_cache`.  It downloads an updated version of the
information stored in `abs_cachelist`---`abs_cachelist` is simply a saved return
of `abs_cache(lang = "en")`.  Updated cache information can be used in the
`abs_search` and `abs_stats` functions by setting the `cache` parameter to the
saved results returned from `abs_cache`. It is advisable to use the updated
cache information to ensure that you have access to the latest available
datasets. Note, however, `abs_cache` can take a long time to download all
datasets, so use sparingly. For this reason, `abs_cache` includes the
`progress=n` argument, which provides an update on progress every `n` datasets.
The second example downloads a new cache in French.^[Note, the [ABS.Stat](http://stat.data.abs.gov.au/) API
currently has limited support for other languages, so data may not be available
in languages other than English.]

```{r eval=FALSE}
# default language is english
new_cache <- abs_cache(progress=10);

new_cache_fr <- abs_cache(lang="fr", progress=10);
```


### Checking dataset dimensions with `abs_dimensions()`

The `abs_dimensions()` functions lists the name of all available dimensions and
the respective dimension type. Typical dimension types are: 'Dimension',
'TimeDimension' and 'Attribute'. 'Dimension' attributes are used in the `filter`
argument of `abs_stats` function. The following example lists the data
dimensions of the 'CPI' dataset.

```{r eval=FALSE}
cpi_dim <- abs_dimensions('CPI');
print(cpi_dim);
```


### Search available data with `abs_search()`

The `abs_search` function essentially has two modes of operation: i) 'dataset
search' mode, and ii) 'indicator search' mode.

In dataset search mode, the function searches and returns datasets matching a
provided regular expression. The following examples demonstrates use of the
function to find ABS datasets relating to the CPI and labour force. The
`code_only` argument specifies whether the function returns all information or
just the matching dataset identifiers.


```{r eval=FALSE}
## Find ABS CPI datasets
abs_search("CPI|consumer price index");
abs_search("CPI|consumer price index", code_only=TRUE);

## Find ABS labour force datasets
abs_search("labour force");
abs_search("^labour force$");
abs_search("^labour force$", code_only=TRUE);
```

In indicator search mode, the function searches through all dimensions of a
specific dataset and returns a list of dimensions and dimension contents
matching all the provided regular expressions. The following example
demonstrates use of the function to find indicators within the CPI data set.

```{r eval=FALSE}
## Find all dimensions in the CPI dataset matching the expression 'All groups'
abs_search("All groups", dataset="CPI");

## Find all dimensions in the CPI dataset matching the expressions 'All groups' and 'Sydney'
abs_search(c("All groups","Sydney"), dataset="CPI");
```

If `code_only=TRUE`, the indicator search function returns only codes for each
matching dimension. This result can be used directly as input to the `filter`
argument of the `abs_stats` function.

```{r eval=FALSE}
## Find all dimensions in the CPI dataset matching the expression 'All groups'
abs_search("All groups", dataset="CPI", code_only=TRUE);

## Find all dimensions in the CPI dataset matching the expressions 'All groups' and 'Sydney'
abs_search(c("All groups","Sydney"), dataset="CPI", code_only=TRUE);
```


### Downloading data with `abs_stats()`

The `abs_stats()` function downloads and returns data from the [ABS.Stat](http://stat.data.abs.gov.au/) API. The
following examples highlight the different ways in which
`abs_stats()` can be used and demonstrate the major optional parameters.

The default value for the `country` parameter is a special value of `all` which
as you might expect, returns data on the selected `indicator` for every
available country or region.

```{r eval=FALSE}
library(raustats)

# Population, total
pop_data <- abs_(indicator = "SP.POP.TOTL", startdate = 2000, enddate = 2002)

head(pop_data)
```

If you are interested in only some subset of countries or regions you can pass
along the specific codes to the `country` parameter. The country and region
codes that can be passed to the `country` parameter correspond to the coded
values from the `iso2c`, `iso3c`, `regionID`, `adminID`, and `incomeID` from the
`countries` data frame in `abs__cachelist` or the return of `abs_cache()`. Any
values from the above columns can mixed together and passed to the same call

```{r eval=FALSE}
library(raustats)

# Population, total
# country values: iso3c, iso2c, regionID, adminID, incomeID
pop_data <- abs_stats(country = c("ABW","AF", "SSF", "ECA", "NOC"),
                      indicator = "SP.POP.TOTL", startdate = 2012, enddate = 2012)

head(pop_data)
```

Queries with multiple indicators return the data in a long data format by
default.

```{r eval=FALSE}
library(raustats)

pop_gdp_long <- abs_stats(country = c("US", "NO"), indicator = c("SP.POP.TOTL", "NY.GDP.MKTP.CD"),
                     startdate = 1971, enddate = 1971)

head(pop_gdp_long)
```

or a wide format if parameter `return_wide = TRUE`. Note that to necessitate
this transformation the `indicator` column is dropped.

```{r eval=FALSE}
library(raustats)

pop_gdp_wide <- abs_stats(country = c("US", "NO"), indicator = c("SP.POP.TOTL", "NY.GDP.MKTP.CD"),
                          startdate = 1971, enddate = 1971, return_wide = TRUE)

head(pop_gdp_wide)
```

### Using `mrv`

If you do not know the latest date an indicator you are interested in is
available for you country you can use the `mrv` instead of `startdate` and
`enddate`. `mrv` stands for most recent value and takes a `integer`
corresponding to the number of most recent values you wish to return

```{r eval=FALSE}
library(raustats)

eg_data <- abs_(country = c("IN"), indicator = 'EG.ELC.ACCS.ZS', mrv = 1)

eg_data
```

You can increase this value and it will return no more than the `mrv`
value. However, if `mrv` is greater than the number of available data it will
return less

```{r eval=FALSE}
library(raustats)

eg_data <- abs_(country = c("IN"), indicator = 'EG.ELC.ACCS.ZS', mrv = 10)

eg_data
```

### Using `gapfill = TRUE`

An additional parameter that can be used along with `mrv` is
`gapfill`. `gapfill` allows you to "fill-in" the values between actual
observations. The "filled-in" value for an otherwise missing date is
the last observed value carried forward.The only difference in the
data call below from the one directly above is `gapfill = TRUE` (the
default is `FALSE`). Note the very important difference

```{r eval=FALSE}
library(raustats)

eg_data <- abs_(country = c("IN"), indicator = 'EG.ELC.ACCS.ZS', mrv = 10, gapfill = TRUE)

eg_data
```

Because `gapfill` returns data that does reflect actual observed
values, use this option with care.

### Using `POSIXct = TRUE`

The default format for the `date` column is not conducive to sorting
or plotting, especially when downloading sub annual data, such as
monthly or quarterly data. To address this, if `TRUE`, the `POSIXct`
parameter adds the additional columns `date_ct` and
`granularity`. `date_ct` converts the default date into a
`POSIXct`. `granularity` denotes the time resolution that the date
represents. This option requires the use of the package `lubridate (>=
1.5.0)`. If `POSIXct = TRUE` and `lubridate (>= 1.5.0)` is not
available, a `warning` is produced and the option is ignored

```{r eval=FALSE}
library(raustats)

oil_data <- abs_(indicator = "CRUDE_BRENT", mrv = 10, freq = "M", POSIXct = TRUE)

head(oil_data)
```

The `POSIXct = TRUE` option makes plotting and sorting dates much easier.

```{r vig-ex-posixct-1, fig.height = 4, fig.width = 7.5, fig.path = "figures/"}
library(raustats)
library(ggplot2)

oil_data <- abs_(indicator = c("CRUDE_DUBAI", "CRUDE_BRENT", "CRUDE_WTI", "CRUDE_PETRO"),
               startdate = "2012M01", enddate = "2014M12", freq = "M", POSIXct = TRUE)

ggplot(oil_data, aes(x = date_ct, y = value, colour = indicator)) + geom_line(size = 1) +
  labs(title = "Crude Oil Price Comparisons", x = "Date", y = "US Dollars")
```


The `POSIXct = TRUE` option also makes plotting time series with
different time coverage seamless.

```{r vig-ex-posixct-2, fig.height = 4, fig.width = 7.5, fig.path = "figures/"}
library(raustats)
library(ggplot2)

# querying seperate for differing time coverage example
gold_data <- abs_(indicator = "GOLD", mrv = 120, freq = "M", POSIXct = TRUE)
plat_data <- abs_(indicator = "PLATINUM", mrv = 60, freq = "M", POSIXct = TRUE)

metal_data <- rbind(gold_data, plat_data)

ggplot(metal_data, aes(x = date_ct, y = value, colour = indicator)) + geom_line(size = 1) +
  labs(title = "Precious Metal Prices", x = "Date", y = "US Dollars")
```


# RBA statistics access functions

### Finding available RBA data tables with `rba_table_cache`

The `rba_table_cache` function returns a dataset of all available RBA
statistical tables. The function scans the RBA website and returns a list of all
[Statistical tables](https://www.rba.gov.au/statistics/tables/), [Historical
data tables](https://www.rba.gov.au/statistics/historical-data.html) and
[Discontinued data
tables](https://www.rba.gov.au/statistics/discontinued-data.html). The dataset
has four columns:
  * table_code
  * table_name
  * table_type -- one of either statistical table, historical data or
    discontinued data.
  * url.

The `rba_table_cache` function has no arguments. The following example shows use
of the function and the returned output.

```{r eval=FALSE}
rba_cache <- rba_table_cache()

head(rba_cache)
```


### Search available data with `rba_search()`

The `rba_search` function searches the RBA table cache for statistical
tables matching the regular expression specified in `pattern`.  The
`fields` argument specifies the fields in the RBA table cache to
search---by default the function searches the `table_code` and
`table_name` fields. The `ignore.case` argument allows for case
sensitive regular expression matching---the default is case
insensitive matching. The `cache` argument specifies the RBA table
cache to be used. If omitted, `rba_tablecache` is used.


```{r eval=FALSE}
## Find RBA datasets
rba_search(pattern = "Liabilities and Assets");

```


### Accessing RBA statistical tables with `rba_stats()`

As previously outlined in the Quick Start section, the `rba_stats()`
function provides easy access to RBA statistical tables.  The
`rba_stats()` function has three mutually-exclusive table selection
arguments: `table_no`, `pattern`, and `url`, for selecting RBA
statistical tables.  Sepcifying `table_no` selects tables matching the
specified RBA table number, e.g.  A1, B1, B11.1.  Specifying `pattern`
selects all RBA tables matching the regular expression specified in
`pattern`.  The `url` argument can be used to specify one or more
valid RBA statistical table URLs. 

The function returns a long-format (tidy) table with the following
columns:
  * series_id -- RBA series identifier
  * date -- Date-format date
  * value -- series value
  * title -- data item name
  * description -- data item description
  * frequency -- Frequency (e.g. Annual, Quarterly, Monthly, Weekly, Daily)
  * type -- series type, one of Original, Trend, Seasonally Adjusted
  * units -- unit of measure (e.g. Percent, $ Millions, Index Numbers, Proportion)
  * source -- data source organisation
  * data_type -- data type (e.g. Original, Derived)
  * table_no -- RBA statistics table number (if available)
  * table_title -- RBA statistics table title.


The following examples demonstrate
typical uses of the function and the various function arguments. The
first example downloads RBA Statistical Table A1 -- Liabilities and
Assets using `table_no`.

```{r eval=FALSE}
# Latest quarterly national accounts
rba_a1 <- rba_stats(table_no = "A1")
head(rba_a1)
```

The second example downloads the same tables using the `pattern`
argument to download tables matching the regular expression:
'Liabilities and Assets.+A1'. 

```{r eval=FALSE}
# Latest quarterly national accounts
rba_a1 <- rba_stats(pattern = "Liabilities and Assets.+A1")
head(rba_a1)
```

And the third example downloads the same tables using the relevant
URLs. The example presented below first returns a list of all RBA
tables matching the supplied regular expression ('Liabilities and
Assets.+A1') and then uses the returned URLs to download each data
set.

```{r eval=FALSE}
# Latest quarterly national accounts
a1_tables <- rba_search(pattern = "Liabilities and Assets.+A1")
head(a1_tables)

rba_a1 <- rba_stats(url = a1_tables$url)
head(rba_a1)
```

The `cache` argument specifies the RBA table cache to be used. If
omitted, `rba_tablecache` is used. The `rba_stats` function also
optionally accepts `rba_search` arguments: `fields` and `ignore.case`,
for informing pattern matching. The `cache` argument specifies the RBA
table cache to be used---again, `rba_tablecache` is used if `cache` is
omitted.


### Other RBA helper functions

There are also two RBA helper functions that are called by
`rba_stats`---that download and parse the RBA statistical tables. These additional
functions should not ordinarily need to be called directly, but there
may be situations in which users wish to access these
functions directly. The functions are:

  * `rba_file_download`
  * `rba_read_tss`

The following examples illustrate the use of these functions.

The `rba_download` function downloads and saves RBA tables from a
supplied URL, and returns a character vector of the saved local
filenames. It is called inside the `rba_stats` function and can be
used to directly download one or more RBA statistical table files. It
is most usefully used in conjunction with the `rba_search` function


```{r eval=FALSE}
# Extracted ABS Excel Time Series Spreadsheet
a1_tables <- rba_search(pattern = "Liabilities and Assets.+A1")
downloaded_tables <- rba_file_download(a1_tables$url, exdir=tempdir())

print(downloaded_tables)
```

The `rba_read_tss` function extracts data from standard-format RBA
statistical tables and returns it as a long-format (tidy) data frame.
It is also called within the `rba_stats` function.  The simple example below shows use of the function to read Catalogue
5206.0, Table 1.

```{r eval=FALSE}
# Read specified RBA statistical data spreadsheets
a1_data <- rba_read_tss(downloaded_tables);
head(a1_data)
```

Note, however, that the RBA historical and discontinued data files
formats do not fully conform with statistical tables format, so will
not necessarily be parsed correctly when read with the `rba_read_tss`
function. A separate function to parse these tables will be
developed.


# Getting started with returned data

Statistics returned by the `raustats` package functions are generally in long
(tidy) format data frames. This provides for easy integration and transformation
with packages like `ggplot2`, `tidyr` and `dplyr`. A few examples are provided
here to help users get started.


The following examples highlight how the data downloaded with the
above functions may be used with the above packages. 

<!-- TO BE COMPLETED -->





# Unresolved issues

There are a few behaviours of the
[ABS.Stat](http://stat.data.abs.gov.au/) API (Beta) that may help
explain any potentially unexpected results. These results are known
but no special actions are taken to mitigate them as they are specific
to the API itself. Moreover, as the
[ABS.Stat](http://stat.data.abs.gov.au/) API is, as at the time of
writing, still in Beta and subject to change, some of these issues are
yet to addressed in this version of the `raustats` package.


## Searching in other languages

The [ABS.Stat](http://stat.data.abs.gov.au/) API (Beta version)
appears to include scope to cater to languages other than English. At
the time of writing, French appears to be the only other language
included in the data sets, however the text for all records denoted
with French language attribute are in English. The ABS API calling
functions included in this library allow users to specify their
preferred language, but this functionality has been little tested to
date.


## Query and data limits

As already noted, the [ABS.Stat](http://stat.data.abs.gov.au/) API has
a query string length limit (1000 characters) and record return limit
(1 million records). We have not tested how rigorously these limits
are enforced. Accordingly, the `abs_stats` function includes the
`enforce_api_limits` argument to catch these limits before the query
is submitted. The argument also allows the user to override the limits
and submit the query anyway. This argument may be deprecated in future
versions.

