---
title:  "Introduction to the raustats package"
author: "David Mitchell"
date:   "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{01_-_Using_the_raustats_package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>")
```


# Introduction

The `raustats` package allows researchers to quickly search and download
selected [Australian Bureau of Statistics (ABS)](http://www.abs.gov.au/) and
[Reserve Bank of Australia (RBA)](http://www.rba.gov.au/) data in a programmatic
and reproducible fashion. This facilitates seamless integration of ABS and/or
RBA data into analysis projects, and enables automatic update to the latest
available data.


## Australian Bureau of Statistics

The [Australian Bureau of Statistics (ABS)](http://www.abs.gov.au/) is
Australia’s national statistical agency, providing trusted official statistics
on a wide range of economic, social, population and environmental matters of
importance to Australia. Key ABS statistical collections include:
 
  * Australian National Accounts
  * International Trade
  * Consumer Price Index (CPI)
  * Labour Force
  * Population trends


## Reserve Bank of Australia

The [Reserve Bank of Australia (RBA)](http://www.rba.gov.au/) is Australia's
central bank. In addition to its legislative responsibilities, it collects and
publishes statistics on money, credit, the Australian banking systems and other
relevant economic metrics. Key RBA statistics include:

  * Banking system assets and liabilities
  * Money and credit statistics
  * Household and business finances
  * Interest rates
  * Exchange rates
  * Inflation and inflation expectations.


## ABS & RBA data availability

The ABS and RBA currently make most, or all, of their statistics primarily
available through Excel- and/or CSV-format spreadsheets. The former typically
require some additional re-formatting to produce well-formatted (tidy) data.

The functions in this package help facilitate access to ABS and RBA statistics
through R, returning data in long-format (tidy-like) tables.

The ABS is also developing a data access API:
[ABS.Stat](http://stat.data.abs.gov.au/), which is currently a Beta release and
provides access to a subset of all ABS statistics. This package include
functions to search and download data via the
[ABS.Stat](http://stat.data.abs.gov.au/) API. Please note that as the API is
still in development, changes in the API back-end may break the functions in
this package.


## Main features of the `raustats` package: 

  * Access to all time series and cross-section statistics in the ABS
    statistical catalogue and on the RBA website
  * Access to statistics available through the ABS Beta API:
    [ABS.Stat](http://stat.data.abs.gov.au)
  * Ability to return `POSIXct` dates for easy integration into
    plotting and time-series analysis techniques 
  * Returns data in long (or tidy) format for direct
    integration with packages like `ggplot2`, `tidyr` and `dplyr`
  * Support for regular expression (`grep`) style searching for data
    descriptions and names.



# Quick-start guide

This section provides a quick-start guide on how to download ABS and
RBA statistics.

First, load the library:
```{r message=FALSE, warning=FALSE}
library(raustats)
```

## Downloading ABS Catalogue Statistics

The `abs_cat_stats` function downloads ABS statistics by ABS Catalogue
number. For example, the latest Consumer Price Index (CPI) data set (ABS
Catalogue no. 6401.0) can be downloaded with the following call:

```{r eval=FALSE}
cpi_all <- abs_cat_stats("6401.0")
```

To download the latest CPI data reported in Table 1 only (ABS groups Tables 1
and 2 together), simply supply a regular expression to the `tables` argument:

```{r eval=FALSE}
cpi <- abs_cat_stats("6401.0", tables="Table.+1\\D")
```

Alternatively, `tables` also accepts a regular expression that matches the table
name as specified on the ABS data access page, as follows:

```{r eval=FALSE}
cpi <- abs_cat_stats("6401.0", tables="CPI: All Groups, Index Numbers and Percentage Change")
```

Statistics from previous releases can be accessed using the `releases` argument. A
more detailed explanation of the `abs_cat_stats` function and further examples
are provided [below](#abs-cat-stats-functions).


## Downloading ABS data via [ABS.Stat](http://stat.data.abs.gov.au/)

The `abs_stats` function is used to download ABS statistics via the
[ABS.Stat](http://stat.data.abs.gov.au/) API. For example, to download the
latest CPI data series for 'All Groups' changes in prices across Australia and
each of the eight capital cities, simply call:

```{r eval=FALSE}
cpi_api <- abs_stats("CPI", filter=list(MEASURE=1, REGION=c(1:8,50),
                                        INDEX=10001, TSEST=10, FREQUENCY="Q"))
```

The package includes additional functions to search for datasets and data series
available on [ABS.Stat](http://stat.data.abs.gov.au/). These are documented
further [below](#abs-api-stats-functions).


## Downloading RBA data

The `rba_stats` function downloads statistics available on the RBA website.  For
example, the latest statistics covering the RBA's assets and liabilities (RBA
Statistical Table A1) can be downloaded with the following call:

```{r eval=FALSE}
rba_bs <- rba_stats("A1")
```

A more detailed explanation of `rba_stats` and other RBA data access functions
is provided [below](#rba-stats-functions).



# ABS statistics access functions

##  ABS Catalogue statistics functions {#abs-cat-stats-functions}

The ABS Catalogue statistics functions are split into core functions:

  * `abs_cat_stats`
  * `abs_cat_tables`
  
and several helper functions:

  * `abs_read_tss`
  * `abs_download_data`
  * `abs_unzip_files`


The helper functions are called by the core functions and should generally not
need to be called directly by users---though there are some cases where these
functions may be useful.


### Finding available ABS Catalogue statistics

The ABS does not provide a consolidated searchable list of all current
statistical collections available through the ABS Catalogue. A text search
facility is available on the ABS website:
[www.abs.gov.au](http://www.abs.gov.au), but this package does not currently
provide any functionality to access this facility. Instead, the `abs_cat_cachelist`
data set contained in this package lists the more common ABS Catalogue
statistics.


### Accessing ABS Catalogue statistics with `abs_cat_stats`

As shown in the Quick Start section, the `abs_cat_stats` function provides easy
access to ABS statistics by ABS Catalogue Number.  The following examples
demonstrate typical uses of the function and the various function arguments.

The simplest use of the `abs_cat_stats` function is to download all tables
available in a specified ABS Catalogue series. The Quick Start illustrated how
to download all CPI tables. The following example downloads the latest quarterly
national accounts statistics (ABS Catalogue no. 5206.0):

```{r message=FALSE, results=FALSE}
ana_q <- abs_cat_stats(cat_no = "5206.0")
```

The function returns a long-format (tidy) table with the following
columns:

  * *series_id* -- ABS series identifier
  * *date* -- Date-format date
  * *value* -- observation value
  * *data_item_description* -- data item name and description
  * *series_type* -- series type, generally one of: Original, Trend, Seasonally Adjusted
  * *series_start* -- series start date
  * *series_end* -- series end data
  * *no_obs* -- number of series observations
  * *unit* -- unit type (e.g. Percent, $ Millions, Index Numbers, Proportion)
  * *data_type* -- data type (e.g. Derived)
  * *freq* (frequency) -- Frequency (e.g. Annual, Quarterly, Monthly)
  * *collection_month* -- collection month (integer)
  * *catalogue_no* -- catalogue number (e.g. 5206.0, 6401.0)
  * *publication_title* -- ABS publication title
  * *table_no* -- ABS publication table number (integer)
  * *table_title* -- ABS publication table title.

```{r echo=FALSE, results=TRUE}
## Latest quarterly national accounts
head(ana_q)
```

The `tables` argument allows users to select the set of catalogue tables to be
downloaded by specifying a regular expression to pattern match the ABS table
names, as specified on the ABS web page for the specified Catalogue number---by
default (`tables = "all"`) the function automatically downloads all available
tables from the specified catalogue number. The following sample code downloads
Tables 1 and 2 from Catalogue no. 5206.0.

```{r eval=FALSE, message=FALSE}
ana_q <- abs_cat_stats(cat_no = "5206.0", tables=c("^Table 1\\D", "^Table 2\\D"))
```

The same result may be achieved by specifying one or more regular expressions
matching one or more table names. For example:

```{r eval=FALSE, echo=FALSE}
ana_q <- abs_cat_stats(cat_no = "5206.0",
                       tables=c(".*Key National Accounts Aggregates",
                                ".*Expenditure on Gross Domestic Product (GDP), Chain volume measures"))
```

The `releases` argument enables users to download data from a specified
release. By default, the function downloads the latest available data
(i.e. `releases="Latest"`). The format is a date object or character string
specifying the month and year of release.  For example, the following sample
code downloads Table 1 from the December 2017 release of the quarterly national
accounts. 

```{r eval=FALSE}
ana_2017Q4 <- abs_cat_stats(cat_no="5206.0", tables="Table 1", releases="Dec 2017")
## or
ana_2017Q4 <- abs_cat_stats(cat_no="5206.0", tables="Table 1", releases=as.Date("2017-12-01"))
```

The `releases` argument accepts multiple elements, as per the following example,
which downloads Table 1 from each of the December 2016 and 2017 quarter national accounts:
```{r eval=FALSE}
ana_Q4 <- abs_cat_stats(cat_no="5206.0", tables="Table 1", releases=c("Dec 2017","Dec 2016"))
```

The `abs_cat_stats` function is designed to download both ABS time series
spreadsheets (`types="tss"`) and cross-section spreadsheets
(`types="css"`)---'Data Cubes' in ABS terminology.  ABS time series spreadsheets
generally have a standard format, with a single column for each series, several
header rows containing series metadata and a single row with the unique series
identifier.  ABS cross-section spreadsheet formats vary, depending on the number
of dimensions (categories) available in the data set. These tables typically
have multiple uniquely-identifying header rows.

Presently, the `abs_cat_stats` only includes functionality to download and
process time series spreadsheets---functionality to handle ABS Data Cubes
(cross-section spreadsheets, `types="css"`) is planned to be added in future
versions.

In the meantime, it is possible to download data cubes using a sequence of
`abs_cat_tables`--`abs_cat_download`--`abs_cat_unzip` and piping the result into
a `read_excel` function call. The following example downloads ABS labour force
table: *LM1 - Labour force status by Age, Greater Capital City and Rest of State
(ASGS), Marital status and Sex*:

```{r eval=FALSE}
lf_age <- abs_cat_tables("6291.0.55.001", include_urls=TRUE) %>%
  filter(grepl("LM1.+", item_name)) %>%
  .$path_zip %>%
  abs_cat_download %>%
  abs_cat_unzip %>%
  read_excel(sheet="Data 1", skip=3)
```

<!-- Text when abs_read_css function added

The `abs_cat_stats` can download both time series (`types="tss"`) and
cross-section (`types="css"`) data series.  ABS time series
spreadsheets generally have a standard format, with a single column
for each series, with several headers rows containing series metadata,
including a single row with the unique series identifier.  ABS
cross-section spreadsheet formats vary, depending on the number of
dimensions (categories) available in the data set. These tables
typically have multiple uniquely-identifying header rows. Importing
these tables is handled by specifying `types="css"` and supplying the
header start row (e.g. `header_row=4`, if the series header
information starts in row 4) and the number of header rows
(e.g. `header_rows=2`, if there are two header rows). The following
example shows the use of the function to import cross-section data on
regional household incomes (ABS Catalogue no. xxxx.0).

-- Include examples of downloading css files

-->



### Finding available ABS Catalogue tables with `abs_cat_tables`

The `abs_cat_tables` function returns a list of all tables for one or more
specified ABS Catalogue numbers.  It can be used to identify the relevant
table(s) to download. The following two examples return all available tables for the
latest quarterly national accounts (5206.0) and CPI (6401.0), respectively.

```{r message=FALSE}
ana_tables <- abs_cat_tables(cat_no="5206.0")
```

```{r echo=FALSE}
head(ana_tables)
```

```{r message=FALSE}
## CPI
cpi_tables <- abs_cat_tables(cat_no="6401.0")
```

```{r echo=FALSE}
head(cpi_tables)
```

The `abs_cat_tables` also has three additional arguments: `releases`, `types`
and `include_urls`.  The `releases` argument returns the list of downloadable
tables from the specified release---by default `releases="Latest"`.  Lists of
available tables from earlier releases can be obtained by specifying the month
and year of release, e.g.`releases = "Jun 2017"`.  The `types` argument enables
users to specify which file types to include. Options are 'tss' -- ABS Time
Series Spreadsheets, 'css' -- ABS Data Cubes, and 'pub' -- ABS Publications. The
default is `types = c('tss', 'css')`.  The `include_urls` argument specifies
whether or not to include the URLs of available data tables in the returned
results. The default is `include_urls=FALSE`.

The following example, returns all quarterly national accounts tables in the
September and December 2017 quarter releases, and includes the table URLs.

```{r message=FALSE}
ana_tables <- abs_cat_tables(cat_no="5206.0", releases=c("Sep 2017", "Dec 2017"),
                             include_urls=TRUE)
```

```{r echo=FALSE}
head(ana_tables)
```

And the following example illustrates use of the `abs_cat_tables` to return all
available downloadable Data Cubes for a non-time series collection---the
Australian Statistical Geography Standard (ASGS) main structure classification
and digital boundaries (Catalogue no. 1270.0.55.001).

```{r message=FALSE}
asgs_files <- abs_cat_tables(cat_no="1270.0.55.001", types="css", include_urls=TRUE)
```

```{r echo=FALSE}
head(asgs_files)
```


<!-- Start - New material -->
### Listing all available Catalogue releases with `abs_cat_releases`

The `abs_cat_releases` function returns a table of all available past releases
for a specified ABS Catalogue number.  If `include_urls=TRUE` (default is
`FALSE`), the function includes the full URL for each of the available
releases. The default is `include_urls=FALSE`.  The following examples return
all available releases for the quarterly national accounts (5206.0) and CPI
(6401.0), respectively.

```{r message=FALSE}
ana_releases <- abs_cat_releases(cat_no="5206.0")
```

```{r echo=FALSE}
head(ana_releases)
```

```{r message=FALSE}
## CPI
cpi_releases <- abs_cat_releases(cat_no="6401.0", include_urls=TRUE)
```

```{r echo=FALSE}
head(cpi_releases)
```
<!-- END - New material -->


### Other ABS Catalogue helper functions

As already noted, there are several ABS Catalogue helper functions that are called by
`abs_cat_stats` and `abs_cat_tables`---that download and parse the ABS
Catalogue table files. The main ones are:

  * `abs_cat_download`
  * `abs_cat_unzip`
  * `abs_read_tss`
<!--  * `abs_read_css` -->

The following examples illustrate the use of these functions.

The `abs_cat_download` function downloads and saves ABS Catalogue tables from a
supplied URL. It is called inside the `abs_cat_stats` and can be used directly
to download one or more ABS Catalogue table files. It is most usefully used in
conjunction with the `abs_cat_tables` function, as follows:

```{r message=FALSE, results=FALSE}
tables <- abs_cat_tables("5206.0", releases="Latest", include_urls=TRUE)
downloaded_tables <- abs_cat_download(tables$path_xls, exdir=tempdir())
print(downloaded_tables)
```

The `abs_cat_unzip` function extracts Excel files from compressed ABS zip
archives (see example below). It uses the `utils::unzip` function (using some
standard file locations). There are two arguments: `files` and `exdir` which
have a similar meaning to the `utils::unzip` equivalent arguments. By default
`exdir = tempdir()`.

```{r}
extracted_files <- abs_cat_unzip(downloaded_tables)
print(extracted_files)
```


The `abs_read_tss` function extracts data from standard-formatted ABS Catalogue
time series spreadsheets and returns it as a long-format (tidy) data frame.  The
next example shows use of the function to read Table 1 from the national
accounts (Catalogue 5206.0).

```{r eval=FALSE}
ana_q <- abs_read_tss(extracted_files)
```


##  ABS.Stat statistics access functions {#abs-api-stats-functions}

The `raustats` package also includes a range of functions to list, search and
download data sets and statistics available through
[ABS.Stat](http://stat.data.abs.gov.au/) API. The following subsections outline
the key functions.


### Finding available data with `abs_datasets`

The `abs_datasets` function returns a list of all datasets available through
[ABS.Stat](http://stat.data.abs.gov.au/). The function has two arguments: `lang`
(default is English: `lang="en"`) and `include_notes` (default:
`include_notes=FALSE`). The following example shows the results with notes
included.

```{r message=FALSE, results="asis"}
datasets <- abs_datasets()
head(datasets)
```

#### Cached list of available ABS.Stat datasets `abs_cachelist`

For performance, a cached list of datasets available through the
[ABS.Stat](http://stat.data.abs.gov.au/) API is provided in the `abs_cachelist`
data set included with `raustats`.  `abs_cachelist` is the default source used
in `abs_search()` and `abs_stats()` to find matching ABS datasets. 

By default, `abs_cachelist` is in English. To search indicators in a different
language, you can download an updated copy of `abs_cachelist` using
`abs_datasets()` ans specifying a different language.


<!-- ### Accessing updated available data with `abs_cache()` -->

<!-- Current information about datasets available from the -->
<!-- [ABS.Stat](http://stat.data.abs.gov.au/) API can be obtained by calling -->
<!-- `abs_cache`.^[`abs_cachelist` is simply a saved return of `abs_cache(lang = -->
<!-- "en")`.]  Updated cache information can be used in the `abs_search` and -->
<!-- `abs_stats` functions by setting the `cache` argument equal to the saved results -->
<!-- returned by `abs_cache`. While using updated cache information will ensure you -->
<!-- identify all available datasets, the `abs_cache` function can take a long time -->
<!-- to download all datasets, so use judiciously. Because of this, `abs_cache` -->
<!-- includes functionality to report progress back to the console. The frequency of -->
<!-- reporting can be set with the `progress` argument (by default `progress=10`). -->
<!-- The second example downloads a new cache in French.^[Note, the -->
<!-- [ABS.Stat](http://stat.data.abs.gov.au/) API currently has limited support for -->
<!-- other languages, so data may not be available in languages other than English.] -->

<!-- ```{r eval=FALSE} -->
<!-- ## Default language: English -->
<!-- new_cache <- abs_cache(progress=10) -->
<!-- ## French language version -->
<!-- new_cache_fr <- abs_cache(lang="fr", progress=10) -->
<!-- ``` -->


### Checking dataset dimensions with `abs_dimensions()`

The `abs_dimensions()` functions lists the name of all available dimensions and
the respective dimension type. Typical dimension types are: 'Dimension',
'TimeDimension' and 'Attribute'. 'Dimension' attributes are used in the `filter`
argument of `abs_stats` function. The following example lists the data
dimensions of the 'CPI' dataset.

```{r}
abs_dimensions('CPI')
```


A list of all available dimension codes and descriptions for a particular
dataset can be viewed by selecting the relevant dataset from `abs_cachelist` or
an updated cache list returned by `abs_cache`.

```{r}
str(abs_cachelist[["CPI"]])
```


### Search available data with `abs_search()`

The `abs_search` function essentially has two modes of operation: 

  i) 'dataset search' mode, and
  ii) 'indicator search' mode.

In dataset search mode, the function searches and returns datasets matching the
specified regular expression. The following examples demonstrate use of the
function to find ABS datasets relating to the CPI and labour force. The
`code_only` argument specifies whether the function returns all information or
just the matching dataset identifiers.


```{r}
abs_search("CPI|consumer price index")

abs_search("CPI|consumer price index", code_only=TRUE)
```

```{r}
abs_search("labour force")

abs_search("^labour force$")
```

In indicator search mode, the function searches through all dimensions of a
specific dataset and returns a list of dimensions and dimension contents
matching all the provided regular expressions. The following examples
demonstrates use of the function to find indicators within the CPI data set.

```{r eval=FALSE, echo=FALSE}
abs_search("All groups", dataset="CPI")
```

```{r message=FALSE}
abs_search(c("All groups CPI","Sydney"), dataset="CPI")
```

If `code_only=TRUE`, the indicator search function returns only codes for each
matching dimension. This, in turn, can be used directly as input to the `filter`
argument of the `abs_stats` function. The following two examples returns
dimension codes for datasets matching "All groups CPI" (example 1) and matching
both "All groups CPI" and "Sydney" (example 2).

```{r echo=FALSE, eval=FALSE}
abs_search("All groups CPI", dataset="CPI", code_only=TRUE)
```

```{r message=FALSE}
abs_search(c("All groups CPI","Sydney"), dataset="CPI", code_only=TRUE)
```


### Downloading data with `abs_stats()`

The `abs_stats()` function returns data from specified datasets available via
the [ABS.Stat](http://stat.data.abs.gov.au/) API. The following section outlines
typical use of the `abs_stats()` function, and also describes each of the core
function arguments.

The following example downloads original  All groups CPI index numbers for each
of the eight Australian state and territory capital cities and also the average
for all capital cities.

```{r message=FALSE, results=FALSE}
cpi <- abs_stats(dataset="CPI", filter=list(MEASURE=1, REGION=c(1:8,50),
                                            INDEX=10001, TSEST=10, FREQUENCY="Q"))
```

The filter conditions are:

  * `MEASURE=1` -- 'Index Numbers'
  * `REGION=c(1:8,50)` -- Each of the eight capital cities (1--8) and all eight
    capital cities (50)
  * `INDEX=10001` -- 'All groups CPI'
  * `TSEST=10` -- 'Original' observations
  * `FREQUENCY=Q` -- Quarterly observations

```{r echo=FALSE}
head(cpi)
```

The `filter` argument can also be set equal to "all", in which case the function
will attempt to download all observations available for the specified
dataset. 


#### ABS.Stat query download constraints

If the data request is large it may breach the
[ABS.Stat](http://stat.data.abs.gov.au/) API query length, record and/or session
time constraints. [ABS.Stat](http://stat.data.abs.gov.au/) has a query string
length limit (maximum of 1000 characters) on URL queries, a record return limit
(1 million records) and session time limits (maximum 10-minute session time
limit).^[See the [ABS.Stat Web Services User Guide
FAQ](http://www.abs.gov.au/ausstats/abs@.nsf/Lookup/by%20Subject/1407.0.55.002~User%20Guide~Main%20Features~Frequently%20Asked%20Questions~9)
for further details.] Queries that breach these limits will need to be
re-specified as multiple separate calls to obtain the required data.

For example, the following `abs_stats` function call, attempts to download all
series available for the CPI dataset, but the specified query length (1191
characters) exceeds maximum request URL character limit.

```{r eval=FALSE}
cpi <- abs_stats(dataset="CPI", filter="all")
```

By default, `abs_stats` checks whether the query string length and the estimated
number of records to be returned and will halt execution if the query breaches
any of these conditions. Setting the `enforce_api_limits = FALSE` (default:
`TRUE`) will ignore these checks and submit the query to the
[ABS.Stat](http://stat.data.abs.gov.au/) API anyway---though this is not
recommended. 
<!-- If the query fails it will return an error. For example, the following -->
<!-- function call will return an error. -->

```{r eval=FALSE}
cpi <- abs_stats(dataset="CPI", filter="all", enforce_api_limits=FALSE)
```
<!-- #> lexical error: invalid char in json text. -->
<!-- #>                                        http://stat.data.abs.gov.au/SDM -->
<!-- #>                      (right here) ------^ -->


Setting the `return_url = TRUE` (default: `FALSE`) will return the RESTful URL
query string, but does not submit the query to the
[ABS.Stat](http://stat.data.abs.gov.au/) API, see the following example function
call and output.

```{r echo=TRUE}
abs_stats(dataset="CPI", filter=list(MEASURE=1, REGION=c(1:8,50),
                                     INDEX=10001, TSEST=10, FREQUENCY="Q"),
          return_url=TRUE)
```

The `abs_search` function can be used to specify the filter. For example, the
following code block produces the same filter list, specified in the previous
example, and can subsequently be supplied to the `abs_stats` `filter` argument.


```{r message=FALSE, results=FALSE}
filter_lst <- abs_search(c("Index numbers", "All groups",
                           "Sydney|Melbourne|Brisbane|Adelaide|Perth|Hobart|Darwin|Canberra|capital cities",
                           "Original", "Quarterly"),
                         dataset="CPI", code_only=TRUE)
cpi <- abs_stats("CPI", filter = filter_lst)
```

#### Splitting an ABS filter list into multiple short queries

The `purrr` package functions: `map` and `cross` can be used to split a long
query into a list of several shorter length queries to submit to
[ABS.Stat](http://stat.data.abs.gov.au/). For example, the following code block
first uses `abs_search` to create a filter list to download estimated resident
population (ERP), by state/territory, age (year) and sex, for June 2017, and
splits the filter into groups no longer than 26 elements in length (`purrr::map`
and `split`), and combines them into complete list (`purrr::cross`). The final
step imports the data by iterating over the filter list (`ds_filter`) using
`lapply` and combining the results into a data frame using `bind_rows`.

```{r message=FALSE, eval=FALSE, results=FALSE}
## ERP dataset ID
abs_ds <- abs_search(pattern="quarterly.*population.*estimates") %>%
  select(id) %>% unlist;

## Create filter 
ds_filter <- abs_search(pattern=c("Estimated Resident Population", "Males|Females|Persons",
                                  "^\\d{1,3}$", "Jun-2017"),
                        dataset=abs_ds, code_only=TRUE) %>%
  map(~ .x %>% split(., ceiling(seq_along(.)/26))) %>%
  cross;
  
## Download Jun 2017 ERP
erp_st_age_sex_2017 <-
  lapply(ds_filter,
         function(i_filter)
           abs_stats(abs_ds,
                     start_date="2017-Q2", end_date="2017-Q2",
                     filter=i_filter)
         ) %>%
  bind_rows;
```


#### Other ABS.Stat query arguments

Users can also limit the date range to return by specifying one or
both`start_date` and `end_date` arguments. These accept either a numeric or
character arguments---if numeric it must be a four-digit year. If a character
string it can be either a monthly, quarterly, half-year or financial year as
formatted as follows: month -- '2016-M01', quarter -- '2016-Q1', half-year --
'2016-B2', financial year -- '2016-17'.  The following example returns all CPI
observations between September 2015 and June 2018.

```{r message=FALSE, results=FALSE}
cpi <- abs_stats(dataset="CPI", filter=filter_lst,
                 start_date = "2015-Q3", end_date = "2018-Q2")
```

The other arguments `dimensionAtObservation` and `detail` provide refinements to
the URL query. These need not be modified by the user---the defaults are:
`dimensionAtObservation='AllDimensions` and `detail='Full'`.


# RBA statistics access functions {#rba-stats-functions}

### Finding available RBA data tables with `rba_table_cache`

The `rba_table_cache` function returns a dataset of all available RBA
statistical tables. The function scans the RBA website and returns a list of all
[Statistical tables](https://www.rba.gov.au/statistics/tables/), [Historical
data tables](https://www.rba.gov.au/statistics/historical-data.html) and
[Discontinued data
tables](https://www.rba.gov.au/statistics/discontinued-data.html). (The
`rba_cachelist` data set included in this package contains a pre-saved list of
all available RBA statistical tables.)

The dataset has four columns:

  * table_code
  * table_name
  * table_type -- one of either statistical table, historical data or
    discontinued data.
  * url.

The `rba_table_cache` function has no arguments. The following example shows use
of the function and the returned output.

```{r}
rba_cache <- rba_table_cache()
```

```{r echo=FALSE}
head(rba_cache)
```

### Search available data with `rba_search()`

The `rba_search` function scans the RBA table cache for statistical tables
matching a regular expression supplied to `pattern`.  The `fields` argument
specifies the fields in the RBA table cache to search---by default the function
searches the `table_code` and `table_name` fields only. The `ignore.case`
argument allows for case sensitive regular expression matching---the default is
case insensitive matching (`ignore.case=TRUE`). The `update_cache` argument
specifies whether the function uses the list of RBA tables supplied with the
package `rba_cachelist` (`update_cache=FALSE`, the default) or to update list of
tables from the RBA website (`update_cache=TRUE`).


```{r}
rba_search(pattern = "Liabilities and Assets")
```


### Accessing RBA statistical tables with `rba_stats()`

As previously outlined in the Quick Start section, the `rba_stats()` function
provides easy access to RBA statistical tables.  The `rba_stats()` function has
three mutually-exclusive table selection arguments: `table_no`, `pattern`, and
`url`, for selecting RBA statistical tables.  Specifying `table_no` selects
tables matching the specified RBA table number, e.g.  A1, B1, B11.1.  Specifying
`pattern` selects all RBA tables matching the regular expression specified in
`pattern`.  The `url` argument can be used to specify one or more valid RBA
statistical table URLs.

The function returns a long-format (tidy) table with the following columns:

  * *series_id* -- RBA series identifier
  * *date* -- Date-format date
  * *value* -- observation value
  * *title* -- data item name
  * *description* -- data item description
  * *frequency* -- Frequency (e.g. Annual, Quarterly, Monthly, Weekly, Daily)
  * *type* -- series type, one of Original, Trend, Seasonally Adjusted
  * *units* -- unit of measure (e.g. Percent, $ Millions, Index Numbers, Proportion)
  * *source* -- data source organisation
  * *data_type* -- data type (e.g. Original, Derived)
  * *table_no* -- RBA statistics table number (if available)
  * *table_title* -- RBA statistics table title.


The following examples demonstrate typical uses of the function and the various
function arguments. The first example downloads RBA Statistical Table A1 --
Liabilities and Assets using `table_no`.

```{r message=FALSE, results=FALSE}
rba_a1 <- rba_stats(table_no = "A1")
```

```{r echo=FALSE}
head(rba_a1)
```

The second example downloads the same tables using the `pattern` argument to
download tables matching the regular expression: 'Liabilities and Assets.+A1'.

```{r eval=FALSE}
rba_a1 <- rba_stats(pattern = "Liabilities and Assets.+Summary")
```

And the third example downloads the same tables using the relevant URLs. The
example presented below first returns a list of all RBA tables matching the
supplied regular expression ('Liabilities and Assets.+A1') and then uses the
returned URLs to download each data set.

```{r message=FALSE, results=FALSE}
a1_tables <- rba_search(pattern = "Liabilities and Assets.+Summary")
rba_a1 <- rba_stats(url = a1_tables$url)
```

```{r echo=FALSE}
head(rba_a1)
```

Again, the `update_cache` argument specifies whether the function uses the list
of RBA tables supplied with the package with the package `rba_cachelist` (the
default) or to update list of tables from the RBA website. The `rba_stats`
function also optionally accepts `rba_search` arguments: `fields` and
`ignore.case`, for informing pattern matching.

At present, the `rba_stats` function only handles [Statistical
tables](https://www.rba.gov.au/statistics/tables/), which have a consistently
structured format comprising full metadata and observations. Functionality to
handle [Historical data
tables](https://www.rba.gov.au/statistics/historical-data.html) and
[Discontinued data
tables](https://www.rba.gov.au/statistics/discontinued-data.html) are yet to be
added.


### Other RBA helper functions

There are also two RBA helper functions that are called by
`rba_stats`---that download and parse the RBA statistical
tables. These additional functions should not ordinarily need to be
called directly, but there may be situations in which users wish to
access these functions directly. The functions are:

  * `rba_file_download`
  * `rba_read_tss`

The following examples illustrate the use of these functions.

The `rba_download` function downloads and saves RBA tables from a supplied URL,
and returns a character vector of the saved local filenames. It is called inside
the `rba_stats` function and can be used to directly download one or more RBA
statistical table files. It is most usefully used in conjunction with the
`rba_search` function.


```{r message=FALSE, results=FALSE}
a1_tables <- rba_search(pattern = "Liabilities and Assets.+Summary")
downloaded_tables <- rba_file_download(a1_tables$url, exdir=tempdir())
print(downloaded_tables)
```

The `rba_read_tss` function extracts data from standard-format RBA statistical
tables and returns it as a long-format (tidy) data frame.  It is also called by
the `rba_stats` function.  The following call will extract data from the
previously downloaded tables.

```{r eval=FALSE}
a1_data <- rba_read_tss(downloaded_tables)
```

```{r echo=FALSE, eval=FALSE}
head(a1_data)
```


# Using data returned by `raustats`

Statistics returned by the `raustats` package functions are generally in long
(tidy) format data frames. This provides for easy integration with packages like
`ggplot2`, `tidyr` and `dplyr`. The following example illustrates how data
downloaded using the `raustats` can be easily transformed and plotted. This
example uses data from ABS' *Private New Capital Expenditure and Expected
Expenditure* (ABS Catalogue no. 5265.0).

First download selected tables from Catalogue no. 5265.0.
```{r message=FALSE, results=FALSE}
capex_q <-
  abs_cat_stats("5625.0",
                tables=c("Actual Expenditure by Type of Asset and Industry - Current Prices",
                         "Actual Expenditure, By Type of Industry - Chain Volume Measures",
                         "Actual and Expected Capital Expenditure by Industry.+:Current Prices"))
```

Then add a new variable denoting Australian state/territory.

```{r results=FALSE, message=FALSE}
library(dplyr)
## Add state/territory variable
capex_q <- capex_q %>%
  mutate(state = sub(sprintf(".*(%s).*",
                             paste(c("New South Wales","Victoria","Queensland","South Australia",
                                     "Western Australia","Tasmania","Northern Territory",
                                     "Australian Capital Territory","Total \\(State\\)"),
                                   collapse="|")),
                     "\\1", data_item_description, ignore.case=TRUE))
```

Finally, plot quarterly time series mining sector capital expenditure (at
current prices) by Australian state and territory using `ggplot`.

```{r fig.height=4, fig.width=7, fig.path="figures/", error=TRUE, message=TRUE, verbose=TRUE}
library(ggplot2)
## Filter mining capital expenditure
capex_q_min <- capex_q %>%
  filter(grepl("mining", data_item_description, ignore.case=TRUE)) %>%
  filter(grepl("actual", data_item_description, ignore.case=TRUE)) %>%
  filter(grepl("current price", data_item_description, ignore.case=TRUE)) %>%
  filter(grepl("Total \\(Type of Asset.+\\)", data_item_description, ignore.case=TRUE))

ggplot(data=capex_q_min) +
  geom_line(aes(x=date, y=value/10^3, colour=state)) +
  scale_x_date(date_labels="%b\n%Y") +
  scale_y_continuous(limits=c(0, NA)) +
  labs(title="Australian mining sector capital expenditure, by state",
       y="Capital expenditure ($ billion)", x=NULL) +
  guides(colour = guide_legend(title=NULL)) + 
  theme(plot.title = element_text(hjust=0.5),
        legend.box = "horizontal",
        legend.position = "bottom",
        axis.text.x=element_text(angle=0, size=8))
```


# Unresolved issues

There are a few behaviours of the [ABS.Stat](http://stat.data.abs.gov.au/) API
that may help explain any unexpected results. As the
[ABS.Stat](http://stat.data.abs.gov.au/) API is still in Beta release and
subject to revision, some of these issues will be addressed in future versions
of the `raustats` package, as [ABS.Stat](http://stat.data.abs.gov.au/)
transitions towards a stable release version.


## Searching in other languages

The [ABS.Stat](http://stat.data.abs.gov.au/) API (Beta version) includes scope
to cater to multiple languages. However, at the time of writing, French appears
to be the only other language included in the data sets. Moreover, the text for
many records that are denoted as French appear to be in English. The ABS API
calling functions included in this library allow users to specify their
preferred language, but this functionality has been little tested to date.


## Query, data and session constraints

As already noted, the [ABS.Stat](http://stat.data.abs.gov.au/) API has a query
string length limit (1000 characters) and record return limit (1 million
records). We have not tested how rigorously these limits are
enforced. Accordingly, the `abs_stats` function includes the
`enforce_api_limits` argument to catch these limits before the query is
submitted. The argument also allows the user to override the limits and submit
the query anyway. This argument may be deprecated in future versions.

The [ABS.Stat](http://stat.data.abs.gov.au/) API also has a 10-minute session
time limit for users to download datasets via the SDMX service. The functions in
`raustats` almost exclusively use the SDMX-JSON query interface, so it is not
clear whether the session time limit applies. If time limits are an issue, the ABS
advises users to submit multiple smaller requests to retrieve the required data.


# Resources

  * [ABS.Stat -- http://stat.data.abs.gov.au/](http://stat.data.abs.gov.au/)
  * [ABS.Stat FAQ](http://stat.data.abs.gov.au/)
  
