---
title:  "Introduction to `raustats`"
author: "David Mitchell"
date:   "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{01_-_Using_the_raustats_package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>")
```


# Introduction

The `raustats` package provides functions to quickly search and download
selected [Australian Bureau of Statistics (ABS)](http://www.abs.gov.au/) and
[Reserve Bank of Australia (RBA)](http://www.rba.gov.au/) data sets in a
programmatic and reproducible fashion, facilitating integration of the latest
ABS and RBA data in analysis.

## Australian Bureau of Statistics

The [Australian Bureau of Statistics (ABS)](http://www.abs.gov.au/) is
Australia’s national statistical agency, providing trusted official statistics
on a wide range of economic, social, population and environmental matters of
importance to Australia. Key ABS statistical collections include:
 
  * Australian National Accounts
  * International Trade
  * Consumer Price Index (CPI)
  * Labour Force
  * Population trends


### Changes to ABS URLs

This version of `raustats` incorporates changes in the ABS URL structures
introduced in September 2020 to make the URL structures more predictable and
user friendly.^[See
[www.abs.gov.au/information-data-users-webscrapers-and-developers](https://www.abs.gov.au/information-data-users-webscrapers-and-developers).]
The ABS continues to support the previous catalogue number based URLs. All
previous functions have been updated and now include two arguments for accessing
ABS data sets--- by `title` or `cat_no`.


## Reserve Bank of Australia

The [Reserve Bank of Australia (RBA)](http://www.rba.gov.au/) is Australia's
central bank. In addition to its legislative responsibilities, it collects and
publishes statistics on money, credit, the Australian banking systems and other
relevant economic metrics. Key RBA statistics include:

  * Banking system assets and liabilities
  * Money and credit statistics
  * Household and business finances
  * Interest rates
  * Exchange rates
  * Inflation and inflation expectations.



## ABS & RBA data access

The ABS and RBA currently publish the majority of their statistics in Excel
spreadsheets or CSV-format files. The functions in this package help facilitate
access to ABS and RBA statistics through R, returning data in tidy tables.

ABS also makes some of its data available via API:
[ABS.Stat](http://stat.data.abs.gov.au/).  This package also includes functions
to access to search and download data via
[ABS.Stat](http://stat.data.abs.gov.au/). The ABS notes that
[ABS.Stat](http://stat.data.abs.gov.au/) is currently a Beta release and that it
may introduce changes at any time that may break the functions in this
package.^[See
[www.abs.gov.au/information-data-users-webscrapers-and-developers](https://www.abs.gov.au/information-data-users-webscrapers-and-developers).]


## Main features of the `raustats` package: 

* Access to all time series statistics in the ABS statistical catalogue and on
  the RBA website
* Access to statistics available through [ABS.Stat](http://stat.data.abs.gov.au)
* Returns `POSIXct` format dates for easy integration into plotting and
  time-series analysis techniques
* Returns data in long (or tidy) format for direct integration with packages
  like `ggplot2`, `tidyr` and `dplyr`
* Support for regular expression (`grep`) style searching for data descriptions
  and names.



# Quick-start guide

This section provides a quick-start guide on how to download ABS and
RBA statistics.

First, load the library:
```{r message=FALSE, warning=FALSE}
library(raustats)
```

## Downloading ABS Catalogue Statistics

The `abs_cat_stats` function downloads ABS statistics either by series title or
by the previously used ABS Catalogue number. For example, the latest Consumer
Price Index (CPI) data spreadsheets (previous ABS Catalogue no. 6401.0) can be
downloaded with either of the following calls:

```{r eval=FALSE}
cpi_all <- abs_cat_stats("Consumer Price Index, Australia");
cpi_all <- abs_cat_stats(cat_no="6401.0")
```

The `tables` argument can be used to supply a regular expression to download a
subset of all available tables. For example, the following call downloads only
Table 1:

```{r eval=FALSE}
cpi <- abs_cat_stats("Consumer Price Index, Australia", tables="Table.+1\\D")
```

Alternatively, `tables` also accepts a regular expression that matches the table
name as specified on the ABS data access page, as follows:

```{r eval=FALSE}
cpi <- abs_cat_stats("Consumer Price Index, Australia",
                     tables="CPI: All Groups, Index Numbers and Percentage Change")
```

Statistics from previous releases can be accessed using the `releases`
argument. A more detailed explanation of the `abs_cat_stats` function and
further examples are provided [below](#abs-cat-stats-functions).


## Downloading ABS data via [ABS.Stat](http://stat.data.abs.gov.au/)

The `abs_stats` function downloads ABS statistics from 
[ABS.Stat](http://stat.data.abs.gov.au/). For example, to download the latest
CPI data series for 'All Groups' change in prices across Australia and each of
the eight capital cities:

```{r eval=FALSE}
cpi_api <- abs_stats("CPI", filter=list(MEASURE=1, REGION=c(1:8,50),
                                        INDEX=10001, TSEST=10, FREQUENCY="Q"))
```

The package includes additional functions to search for datasets and data series
available on [ABS.Stat](http://stat.data.abs.gov.au/), which are documented
[below](#abs-api-stats-functions).



## Downloading RBA data

The `rba_stats` function downloads statistics available on the RBA website.  For
example, the latest statistics covering the RBA's assets and liabilities (RBA
Statistical Table A1) can be downloaded with the following call:

```{r eval=FALSE}
rba_balsheet <- rba_stats("A1")
```

A more detailed explanation of `rba_stats` and other RBA data access functions
is provided [below](#rba-stats-functions).



# ABS data access functions

##  ABS Catalogue statistics functions {#abs-cat-stats-functions}

`raustats` provides five core ABS Catalogue statistics access functions:

  * `abs_cat_search`
  * `abs_cat_series`
  * `abs_cat_releases`
  * `abs_cat_tables`
  * `abs_cat_stats`

and several helper functions:

  * `abs_read_tss`
  * `abs_download_data`
  * `abs_unzip_files`

The helper functions are called by the core functions and will generally not
need to be called directly. However, there are cases where access to these
functions may be handy.



### A typical ABS data access workflow

A typical workflow might involve calling each of the core functions in turn.

1. The `abs_cat_search()` and `abs_cat_series()` functions may be used to
   identify the formal series title. The following examples illustrate use of
   the `abs_cat_search` and `abs_cat_series` functions, respectively, to
   identify ABS National Accounts, National Income and Expenditure publication
   title.

   The first term 
```{r eval=FALSE}
sel_search <- abs_cat_search(pattern="national.*accounts", sort_by='Newest') %>%
  dplyr::filter(grepl("National.+Income.+Expenditure", title, ignore.case=TRUE)) %>%
  dplyr::slice_head(n=1);
```

```{r eval=FALSE}
sel_title <- abs_cat_series(pattern="national.*accounts") %>%
  filter(grepl("National.+Income.+Expenditure", title, ignore.case=TRUE)) %>%
  dplyr::slice_head(n=1);
```

2. The `abs_cat_releases()` function can then be used to return the set of
   available releases.
   
```{r eval=FALSE}
sel_releases <- abs_cat_releases(sel_title$topic[1]) %>%
  filter(grepl("National.+Income.+Expenditure", title, ignore.case=TRUE))
```
   
3. The `abs_cat_tables()` function can then used to list the set of table for
   specified releases.

```{r eval=FALSE}
sel_tables <- abs_cat_tables(sel_title$topic, releases="Latest");
```

4. Finally, `abs_cat_stats()` function is used to download selected tables --
   the following example downloads Tables 1 and 2.

```{r eval=FALSE}
sel_data <- abs_cat_stats(sel_title$topic, releases="Latest", tables="Table\\s*(1|2)\\.");
```

Further examples of use of the core functions are provided next.


### Searching the ABS website

The ABS website provides text search facility that can be accessed via the
`abs_cat_search()` function. The function supports the _simple_ and _advanced_
search capabilities provided by the ABS website, including the ability to search
contents using 'Any' words, or 'All' words, an exact 'Phrase' and/or by
exclusion ('Not'). The function also supports filters to refine the search
period and sort the results by relevance, title or date. The date filter can be
specified by date or text expression. Text date range expressions can include:
'Today', 'Past week', 'Past fortnight', 'Past month', 'Past 3 months', 'Past 6
months' and 'Past year'.

The following example searches the ABS site for occurrences of the words: 'gross',
'domestic' and 'product' and returns results published within the past 3 months and
returns all relevant results in a five-column data frame that 
The function returns a long-format (tidy) table with columns: 

  * *title* -- ABS series title
  * *reference_period* -- ABS publication reference period
  * *release_date* -- publication release date
  * *url* -- ABS web page
  * *resource* -- resource type, generally one of either 'Statistical analysis
    and data' or 'Article'.

The title field can be used in the other core functions to specify a series title. 

```{r message=FALSE}
results <- abs_cat_search("gross domestic product", date_range="Past 3 months");
```

```{r echo=FALSE}
knitr::kable(results);
```

The ABS advanced search feature enables searching for occurrences of all terms
and within a specified date range. For example, search for all publications with
including the words 'national accounts' released between July 2010 and June
2015:

```{r message=FALSE}
results <- abs_cat_search(list(all="National Accounts"),
                          start_date="2010-07-01", end_date="2015-06-30");
```

Note the function currently limits results returned only to those classified as
either 'Statistical analysis and data' products or 'Article' publications, but
many of the older (archive) releases are not classified and currently excluded
by the `abs_cat_search()` function.

Search results can be sorted by age (`sort_by = 'Newest'` or `sort_by =
'Oldest'`) and alphabetical or order reverse alphabetical order ((`sort_by =
'A-Z'` or `sort_by = 'Z-A'`),

The number of results returned can be refined by adjusting the `n_results` and
`follow_links` arguments, which change the number of results returned per page
and number of page links the function follows when returning results.

Lastly, the search URL submitted by a `abs_cat_search()` call can be viewed by
call the function with `return_url = TRUE`. 


### Finding ABS catalogue series titles

The `abs_cat_series` function is designed to return one or more ABS statistics
titles that match a supplied regular expression. The results can then be used in
other functions. If `include_urls=TRUE` (default is `FALSE`), the function
includes the full URL for each of the available releases. The `level` argument
permits refinement of the search to one or more of ABS statistics 'group',
'view' or 'title' (topic), but it should rarely be needed.  The following
example returns a list of ABS statistical releases matching the regular
expression pattern: "national.*income.*expenditure.*product", and includes the
publication URLs.^[Note that the `abs_cat_series` function, scans the ABS
website for all statistical collections upon first use each session.]

```{r message=FALSE}
abs_title <- abs_cat_series(pattern="national.*income.*expenditure.*product",
                            include_urls=TRUE);
```

```{r echo=FALSE}
knitr::kable(abs_title)
```


### List all available ABS releases

The `abs_cat_releases` function returns a table of all current, previous and
archived statistical releases for a specified ABS statistical series. ABS
statistical series can be specified by `title` or by previous catalogue number
(`cat_no`).^[The ABS website handles redirection of series specified by
catalogue number.] The function returns a data frame listing the publication
title (title), publication type (Latest release, Previous release or Archive
release) and reference period. If `include_urls=TRUE` (default is `FALSE`), the
function includes the full URL for each available release. The following
examples return all available releases for the quarterly national accounts and
the consumer price index (CPI), by title and catalogue number respectively.

```{r message=FALSE}
ana_releases <- abs_cat_releases("Australian National Accounts: National Income, Expenditure and Product")
```

```{r message=FALSE, eval=FALSE}
ana_releases <- abs_cat_releases(cat_no="5206.0", include_urls=TRUE);
```

```{r echo=FALSE}
knitr::kable(ana_releases %>% dplyr::slice_head(n=10))
```

```{r message=FALSE}
cpi_releases <- abs_cat_releases(title="Consumer Price Index, Australia", include_urls=TRUE)
```

```{r message=FALSE, eval=FALSE}
cpi_releases <- abs_cat_releases(cat_no="6401.0", include_urls=TRUE)
```

```{r echo=FALSE}
knitr::kable(cpi_releases %>% dplyr::slice_head(n=10))
```


### Finding available ABS statistical tables

The `abs_cat_tables` function returns a list of all tables available for one or
more specified ABS statistical series and specified release. It can be used to
identify the relevant table(s) to download. The following two examples return a
list of all available tables for the latest quarterly national accounts and CPI,
respectively.

```{r message=FALSE, eval=FALSE}
ana_tables <- abs_cat_tables(title="Australian National Accounts: National Income, Expenditure and Product",
                             include_urls=TRUE);
```

```{r message=FALSE}
ana_tables <- abs_cat_tables(cat_no="5206.0");
```

```{r echo=FALSE}
knitr::kable(ana_tables %>% head(10));
```

```{r message=FALSE}
## CPI
cpi_tables <- abs_cat_tables("Consumer Price Index, Australia",  releases="Latest")
```

```{r echo=FALSE}
knitr::kable(ana_tables %>% head(10));
```

The `releases` argument returns the list of downloadable tables for the
specified release(s)---by default `releases="Latest"`.  Lists of available
tables from earlier releases can be obtained by specifying the reference period (`reference_period`)
returned by the `abs_cat_releases()` function, e.g. "December 2019" or "Jun
2017". The `include_urls` argument specifies whether or not to include the URLs
of all available data tables in the returned results (the default is `FALSE`).
The following example, returns all quarterly national accounts tables in the
September and December 2017 quarter releases, and includes the table URLs.

```{r message=FALSE}
ana_tables <-
  abs_cat_tables("Australian National Accounts: National Income, Expenditure and Product",
                 releases=c("Sep 2017", "Dec 2017"))
```

And the following example illustrates use of the `abs_cat_tables` to return all
available downloadable Data Cubes for a non-time series collection---the
Australian Statistical Geography Standard (ASGS) main structure classification
and digital boundaries (Catalogue no. 1270.0.55.001).

```{r message=FALSE, eval=FALSE}
asgs_files <- abs_cat_tables(cat_no="1270.0.55.001", include_urls=TRUE)
```

```{r echo=FALSE, eval=FALSE}
head(asgs_files)
```



### Downloading ABS catalogue statistics

Finally, the `abs_cat_stats` function can be used to download ABS statistics
from one or more ABS statistical series tables.  The following examples
demonstrate typical uses of the function.

The simplest use of the `abs_cat_stats` function is to download all tables
available in a specified ABS Catalogue series. The Quick Start illustrated how
to download all CPI tables. The following example downloads the latest quarterly
national accounts statistics:

```{r message=FALSE, results=FALSE}
ana_q <- abs_cat_stats("Australian National Accounts: National Income, Expenditure and Product")
```

The function returns a long-format (tidy) table with the following columns:

  * *series_id* -- ABS series identifier
  * *date* -- Date-format date
  * *value* -- observation value
  * *data_item_description* -- data item name and description
  * *series_type* -- series type, generally one of: Original, Trend, Seasonally Adjusted
  * *series_start* -- series start date
  * *series_end* -- series end data
  * *no_obs* -- number of series observations
  * *unit* -- unit type (e.g. Percent, $ Millions, Index Numbers, Proportion)
  * *data_type* -- data type (e.g. Derived)
  * *freq* (frequency) -- Frequency (e.g. Annual, Quarterly, Monthly)
  * *collection_month* -- collection month (integer)
  * *catalogue_no* -- catalogue number (e.g. 5206.0, 6401.0)
  * *publication_title* -- ABS publication title
  * *table_no* -- ABS publication table number (integer)
  * *table_title* -- ABS publication table title.


```{r echo=FALSE, results=TRUE}
## Latest quarterly national accounts
knitr::kable(ana_q %>% head(10))
```

The `tables` argument allows users to select the set of data tables to be
downloaded by specifying a regular expression to pattern match the ABS table
names, as specified on the ABS web page for the specified statistical series---by
default (`tables = "all"`) the function automatically downloads all available
tables from the specified series. The following sample code downloads
Tables 1 and 2 from catalogue no. 5206.0.

```{r eval=FALSE, message=FALSE}
ana_q <- abs_cat_stats(cat_no = "5206.0", tables=c("^Table 1\\D", "^Table 2\\D"))
```

The same result may be achieved by specifying one or more regular expressions
matching one or more table names. For example:

```{r eval=FALSE, echo=FALSE}
ana_q <- abs_cat_stats("Australian National Accounts: National Income, Expenditure and Product",
                       tables=c(".*Key National Accounts Aggregates",
                                ".*Expenditure on Gross Domestic Product (GDP), Chain volume measures"));
```

The `releases` argument enables users to download data from a specified
release. By default, the function downloads the latest available data
(i.e. `releases="Latest"`). The format is a date object or character string
specifying the month and year of release.  For example, the following sample
code downloads Table 1 from the December 2017 release of the quarterly national
accounts. 

```{r eval=FALSE}
ana_2017Q4 <- abs_cat_stats("Australian National Accounts: National Income, Expenditure and Product",
                            tables="Table 1\\D", releases="Dec 2017")
```

The `releases` argument accepts multiple elements, as per the following example,
which downloads Table 1 from each of the December 2016 and 2017 quarter national accounts:
```{r eval=FALSE}
ana_q4 <- abs_cat_stats(cat_no="5206.0", tables="Table 1\\D", releases=c("Dec 2017","Dec 2016"))
```

The `abs_cat_stats` function is designed to download ABS time series
spreadsheets only. ABS time series spreadsheets generally have a standard
format, with a single column for each series, several header rows containing
series metadata and a single row with the unique series identifier. By contrast,
ABS cross-section spreadsheet formats vary, depending on the number of
dimensions (categories) available in the data set and cross-section spreadsheets
these tables typically have multiple uniquely-identifying header rows.

The `abs_cat_stats` function only includes functionality to download and process
time series spreadsheets. However, users can download data cube spreadsheets and
import data using a sequence of calls to `abs_cat_tables`, `abs_cat_download`
and `abs_cat_unzip` and piping the result into a `read_excel` function call. For
example, the following command downloads ABS labour force table: *LM1 - Labour
force status by Age, Greater Capital City and Rest of State (ASGS), Marital
status and Sex*:

```{r eval=FALSE}
lf_age <- abs_cat_tables(cat_no="6291.0.55.001", include_urls=TRUE) %>%
  unclass %>% as.data.frame %>%
  dplyr::filter(grepl("LM1.+", table_name)) %>%
  .$file_url %>%
  abs_cat_download %>%
  read_excel(sheet="Data 1", skip=3)
```


### Other ABS Catalogue helper functions

As already noted, `raustats` includes several helper functions, which are called
by the core functions, when downloading and parsing the ABS Catalogue table
files. The main ones are:

  * `abs_cat_download`
  * `abs_cat_unzip`
  * `abs_read_tss`
<!--  * `abs_read_css` -->

The following examples illustrate the use of these functions.

The `abs_cat_download` function downloads and saves ABS Catalogue tables from a
supplied URL. It is called inside the `abs_cat_stats` and can be used directly
to download one or more ABS Catalogue table files. It is most usefully used in
conjunction with the `abs_cat_tables` function, as follows:

```{r message=FALSE, results=FALSE}
tables <- abs_cat_tables("5206.0", releases="Latest", include_urls=TRUE)
downloaded_tables <- abs_cat_download(tables$path_xls, exdir=tempdir())
print(downloaded_tables)
```

The `abs_cat_unzip` function extracts Excel files from compressed ABS zip
archives (see example below). It uses the `utils::unzip` function (using some
standard file locations). There are two arguments: `files` and `exdir` which
have a similar meaning to the `utils::unzip` equivalent arguments. By default
`exdir = tempdir()`.

```{r}
extracted_files <- abs_cat_unzip(downloaded_tables)
print(extracted_files)
```

The `abs_read_tss` function extracts data from standard-formatted ABS Catalogue
time series spreadsheets and returns it as a long-format (tidy) data frame.  The
next example shows use of the function to read Table 1 from the national
accounts (Catalogue 5206.0).

```{r eval=FALSE}
ana_q <- abs_read_tss(extracted_files)
```



##  ABS.Stat statistics access functions {#abs-api-stats-functions}

This section outlines the range of functions available to list, search and
download data sets and statistics available through
[ABS.Stat](http://stat.data.abs.gov.au/). 


### Finding available data with `abs_datasets`

The `abs_datasets` function returns a list of all datasets available through
[ABS.Stat](http://stat.data.abs.gov.au/). The function has two arguments: `lang`
(default is English: `lang="en"`) and `include_notes` (default:
`include_notes=FALSE`). The following example shows the results with notes
included.^[This function is called on first use each session.] 

```{r message=FALSE, results="asis"}
datasets <- abs_datasets()
```

```{r message=FALSE, echo=FALSE, results="asis"}
knitr::kable(datasets %>% head(10))
```


### Checking dataset dimensions with `abs_dimensions()`

The `abs_dimensions()` function lists all available dimensions and dimension
type for a selected dataset available through
[ABS.Stat](http://stat.data.abs.gov.au/). Typical dimensions include: MEASURE,
REGION, INDEX, FREQUENCY, TIME, OBS_STATUS. The following example lists the data
dimensions of the 'CPI' dataset.

```{r warning=FALSE, eval=FALSE}
abs_dimensions('CPI')
```

```{r warning=FALSE, echo=FALSE}
knitr::kable(abs_dimensions('CPI'));
```


### Search available [ABS.Stat](http://stat.data.abs.gov.au/) data

The `abs_search` function essentially has two modes of operation: 

  i) 'dataset search' mode, and
  ii) 'indicator search' mode.

In dataset search mode, the function searches and returns datasets matching the
specified regular expression. The following examples demonstrate use of the
function to find ABS datasets relating to the CPI and labour force. The
`code_only` argument specifies whether the function returns all information or
just the matching dataset identifiers.


```{r}
abs_search("CPI|consumer price index")
abs_search("CPI|consumer price index", code_only=TRUE)
```

```{r}
abs_search("labour force")
```

```{r}
abs_search("^labour force$")
```

In indicator search mode, the function searches through all dimensions of a
specific dataset and returns a list of dimensions and dimension contents
matching all the provided regular expressions. The following examples
demonstrate use of the function to find indicators within the CPI data set.

```{r eval=FALSE, echo=FALSE}
abs_search("All groups", dataset="CPI")
```

```{r message=FALSE}
abs_search(c("All groups CPI","Sydney"), dataset="CPI")
```

If `code_only=TRUE`, the indicator search function returns only codes for each
matching dimension. This, in turn, can be used directly as input to the `filter`
argument of the `abs_stats` function. The following two examples returns
dimension codes for datasets matching "All groups CPI" (example 1) and matching
both "All groups CPI" and "Sydney" (example 2).

```{r echo=FALSE, eval=FALSE}
abs_search("All groups CPI", dataset="CPI", code_only=TRUE)
```

```{r message=FALSE}
abs_search(c("All groups CPI","Sydney"), dataset="CPI", code_only=TRUE)
```


### Downloading [ABS.Stat](http://stat.data.abs.gov.au/) data

The `abs_stats()` function returns data from a specified
[ABS.Stat](http://stat.data.abs.gov.au/) dataset. Typical usage examples are
provided here.

The following example downloads original 'All groups' CPI index numbers for each
of the eight Australian state and territory capital cities and also the average
for all capital cities.

```{r message=FALSE, results=FALSE}
cpi <- abs_stats(dataset="CPI", filter=list(MEASURE=1, REGION=c(1:8,50),
                                            INDEX=10001, TSEST=10, FREQUENCY="Q"))
```

The filter conditions are:

  * `MEASURE=1` -- 'Index Numbers'
  * `REGION=c(1:8,50)` -- Each of the eight capital cities (1--8) and all eight
    capital cities (50)
  * `INDEX=10001` -- 'All groups CPI'
  * `TSEST=10` -- 'Original' observations
  * `FREQUENCY=Q` -- Quarterly observations

```{r echo=FALSE}
head(cpi)
```

The `filter` argument can also be set equal to "all", in which case the function
will attempt to download all observations available for the specified dataset.


#### [ABS.Stat](http://stat.data.abs.gov.au/) query download constraints

If the data request is large it may breach the
[ABS.Stat](http://stat.data.abs.gov.au/) API query length, record and/or session
time constraints. [ABS.Stat](http://stat.data.abs.gov.au/) has a query string
length limit (maximum of 1000 characters) on URL queries, a record return limit
(1 million records) and a maximum 10-minute session time limit).^[See the
[ABS.Stat Web Services User Guide
FAQ](http://www.abs.gov.au/ausstats/abs@.nsf/Lookup/by%20Subject/1407.0.55.002~User%20Guide~Main%20Features~Frequently%20Asked%20Questions~9)
for further details.] Queries that breach these limits will fail with one or
more error/warning messages, depending on the query.

For example, the following `abs_stats` function call fails when it attempts to
download all series available in the CPI dataset, as the specified query length
exceeds the maximum request URL character limit and attempts to return more than
1 million records.

```{r eval=FALSE}
cpi <- abs_stats(dataset="CPI", filter="all")
```

The `abs_stats` function now checks the query string length and if it exceeds
the character length limit, it will automatically separate the query into
multiple separate queries below the length limit, submit them sequentially and
return the combined records. The `abs_stats` function will submit up to 20
sequential queries, but this limit may be increased by increasing the size of
the `requery_limit` argument.

The re-query functionality does not, however, protect against queries returning
more than 1 million records. In those cases, the user will have to filter the
query themselves.  <!-- If the query fails it will return an error. For example,
the following --> <!-- function call will return an error. -->

```{r eval=FALSE}
cpi <- abs_stats(dataset="CPI", filter="all")
```
<!-- #> lexical error: invalid char in json text. -->
<!-- #>                                        http://stat.data.abs.gov.au/SDM -->
<!-- #>                      (right here) ------^ -->


Setting the `return_url = TRUE` (default: `FALSE`) will return the 
[ABS.Stat](http://stat.data.abs.gov.au/) query string and  not submit the
query, for example:

```{r echo=TRUE}
abs_stats(dataset="CPI", filter=list(MEASURE=1, REGION=c(1:8,50),
                                     INDEX=10001, TSEST=10, FREQUENCY="Q"),
          return_url=TRUE)
```

As already noted, the `abs_search` function can be used to specify query filter
settings. For example, the following code block produces the same filter list,
specified in the previous example, and can subsequently be supplied to the
`abs_stats` `filter` argument.


```{r message=FALSE, results=FALSE}
filter_lst <- abs_search(c("Index numbers", "All groups",
                           "Sydney|Melbourne|Brisbane|Adelaide|Perth|Hobart|Darwin|Canberra|capital cities",
                           "Original", "Quarterly"),
                         dataset="CPI", code_only=TRUE)
cpi <- abs_stats("CPI", filter = filter_lst)
```


<!------------------------------------------------------------------------------------------------->
<!-- #### Splitting an ABS filter list into multiple short queries								 -->
<!-- 																							 -->
<!-- The `purrr` package functions: `map` and `cross` can be used to split a long				 -->
<!-- query into a list of several shorter length queries to submit to							 -->
<!-- [ABS.Stat](http://stat.data.abs.gov.au/). For example, the following code block			 -->
<!-- first uses `abs_search` to create a filter list to download estimated resident				 -->
<!-- population (ERP), by state/territory, age (year) and sex, for June 2017, and				 -->
<!-- splits the filter into groups no longer than 26 elements in length (`purrr::map`			 -->
<!-- and `split`), and combines them into complete list (`purrr::cross`). The final				 -->
<!-- step imports the data by iterating over the filter list (`ds_filter`) using				 -->
<!-- `lapply` and combining the results into a data frame using `bind_rows`.					 -->
<!-- 																							 -->
<!-- ```{r message=FALSE, eval=FALSE, results=FALSE}											 -->
<!-- ## ERP dataset ID																			 -->
<!-- abs_ds <- abs_search(pattern="quarterly.*population.*estimates") %>%						 -->
<!--   select(id) %>% unlist;																	 -->
<!-- 																							 -->
<!-- ## Create filter 																			 -->
<!-- ds_filter <- abs_search(pattern=c("Estimated Resident Population", "Males|Females|Persons", -->
<!--                                   "^\\d{1,3}$", "Jun-2017"),								 -->
<!--                         dataset=abs_ds, code_only=TRUE) %>%								 -->
<!--   map(~ .x %>% split(., ceiling(seq_along(.)/26))) %>%										 -->
<!--   cross;																					 -->
<!--   																							 -->
<!-- ## Download Jun 2017 ERP																	 -->
<!-- erp_st_age_sex_2017 <-																		 -->
<!--   lapply(ds_filter,																		 -->
<!--          function(i_filter)																 -->
<!--            abs_stats(abs_ds,																 -->
<!--                      start_date="2017-Q2", end_date="2017-Q2",								 -->
<!--                      filter=i_filter)														 -->
<!--          ) %>%																				 -->
<!--   bind_rows;																				 -->
<!-- ```																						 -->
<!------------------------------------------------------------------------------------------------->


#### Other ABS.Stat query arguments

Users can also limit the date range that will be returned by specifying one or
both `start_date` and `end_date` arguments. These accept either numeric or
character string arguments. Numeric arguments be a four-digit year. Character
strings can be formatted either as monthly, quarterly, half-year or financial
year, as follows: 

* month -- '2016-M01', 
* quarter -- '2016-Q1'
* half-year -- '2016-B2'
* financial year -- '2016-17'. 

The following example returns all CPI observations between September 2015 and
June 2018.

```{r message=FALSE, results=FALSE}
cpi <- abs_stats(dataset="CPI", filter=filter_lst,
                 start_date = "2015-Q3", end_date = "2018-Q2")
```

The remaining arguments to `abs_stats` -- `dimensionAtObservation` and `detail`
-- provide refinements to the URL query. These generally need not be modified by
the user. The defaults are: `dimensionAtObservation='AllDimensions` and
`detail='Full'`.


# RBA statistics access functions {#rba-stats-functions}

### Finding available RBA data tables with `rba_table_cache`

The `rba_table_cache` function returns a dataset of all available RBA
statistical tables. The function scans the RBA website and returns a list of all
[Statistical tables](https://www.rba.gov.au/statistics/tables/), [Historical
data tables](https://www.rba.gov.au/statistics/historical-data.html) and
[Discontinued data
tables](https://www.rba.gov.au/statistics/discontinued-data.html). (The
`rba_cachelist` data set included in this package contains a pre-saved list of
all available RBA statistical tables.)

The dataset has four columns:

  * table_code
  * table_name
  * table_type -- one of either statistical table, historical data or
    discontinued data.
  * url.

The `rba_table_cache` function has no arguments. The following example shows use
of the function and the returned output.

```{r}
rba_cache <- rba_table_cache()
```

```{r echo=FALSE}
head(rba_cache)
```

### Search available data with `rba_search()`

The `rba_search` function scans the RBA table cache for statistical tables
matching a regular expression supplied to `pattern`.  The `fields` argument
specifies the fields in the RBA table cache to search---by default the function
searches the `table_code` and `table_name` fields only. The `ignore.case`
argument allows for case sensitive regular expression matching---the default is
case insensitive matching (`ignore.case=TRUE`). The `update_cache` argument
specifies whether the function uses the list of RBA tables supplied with the
package `rba_cachelist` (`update_cache=FALSE`, the default) or to update list of
tables from the RBA website (`update_cache=TRUE`).


```{r}
rba_search(pattern = "Liabilities and Assets")
```


### Accessing RBA statistical tables with `rba_stats()`

As previously outlined in the Quick Start section, the `rba_stats()` function
provides easy access to RBA statistical tables.  The `rba_stats()` function has
three mutually-exclusive table selection arguments: `table_no`, `pattern`, and
`url`, for selecting RBA statistical tables.  Specifying `table_no` selects
tables matching the specified RBA table number, e.g.  A1, B1, B11.1.  Specifying
`pattern` selects all RBA tables matching the regular expression specified in
`pattern`.  The `url` argument can be used to specify one or more valid RBA
statistical table URLs.

The function returns a long-format (tidy) table with the following columns:

  * *series_id* -- RBA series identifier
  * *date* -- Date-format date
  * *value* -- observation value
  * *title* -- data item name
  * *description* -- data item description
  * *frequency* -- Frequency (e.g. Annual, Quarterly, Monthly, Weekly, Daily)
  * *type* -- series type, one of Original, Trend, Seasonally Adjusted
  * *units* -- unit of measure (e.g. Percent, $ Millions, Index Numbers, Proportion)
  * *source* -- data source organisation
  * *data_type* -- data type (e.g. Original, Derived)
  * *table_no* -- RBA statistics table number (if available)
  * *table_title* -- RBA statistics table title.


The following examples demonstrate typical uses of the function and the various
function arguments. The first example downloads RBA Statistical Table A1 --
Liabilities and Assets using `table_no`.

```{r message=FALSE, results=FALSE}
rba_a1 <- rba_stats(table_no = "A1")
```

```{r echo=FALSE}
head(rba_a1)
```

The second example downloads the same tables using the `pattern` argument to
download tables matching the regular expression: 'Liabilities and Assets.+A1'.

```{r eval=FALSE}
rba_a1 <- rba_stats(pattern = "Liabilities and Assets.+Summary")
```

And the third example downloads the same tables using the relevant URLs. The
example presented below first returns a list of all RBA tables matching the
supplied regular expression ('Liabilities and Assets.+A1') and then uses the
returned URLs to download each data set.

```{r message=FALSE, results=FALSE}
a1_tables <- rba_search(pattern = "Liabilities and Assets.+Summary")
rba_a1 <- rba_stats(url = a1_tables$url)
```

```{r echo=FALSE}
head(rba_a1)
```

Again, the `update_cache` argument specifies whether the function uses the list
of RBA tables supplied with the package with the package `rba_cachelist` (the
default) or to update list of tables from the RBA website. The `rba_stats`
function also optionally accepts `rba_search` arguments: `fields` and
`ignore.case`, for informing pattern matching.

At present, the `rba_stats` function only handles [Statistical
tables](https://www.rba.gov.au/statistics/tables/), which have a consistently
structured format comprising full metadata and observations. Functionality to
handle [Historical data
tables](https://www.rba.gov.au/statistics/historical-data.html) and
[Discontinued data
tables](https://www.rba.gov.au/statistics/discontinued-data.html) are yet to be
added.


### Other RBA helper functions

There are also two RBA helper functions that are called by
`rba_stats`---that download and parse the RBA statistical
tables. These additional functions should not ordinarily need to be
called directly, but there may be situations in which users wish to
access these functions directly. The functions are:

  * `rba_file_download`
  * `rba_read_tss`

The following examples illustrate the use of these functions.

The `rba_download` function downloads and saves RBA tables from a supplied URL,
and returns a character vector of the saved local filenames. It is called inside
the `rba_stats` function and can be used to directly download one or more RBA
statistical table files. It is most usefully used in conjunction with the
`rba_search` function.


```{r message=FALSE, results=FALSE}
a1_tables <- rba_search(pattern = "Liabilities and Assets.+Summary")
downloaded_tables <- rba_file_download(a1_tables$url, exdir=tempdir())
print(downloaded_tables)
```

The `rba_read_tss` function extracts data from standard-format RBA statistical
tables and returns it as a long-format (tidy) data frame.  It is also called by
the `rba_stats` function.  The following call will extract data from the
previously downloaded tables.

```{r eval=FALSE}
a1_data <- rba_read_tss(downloaded_tables)
```

```{r echo=FALSE, eval=FALSE}
head(a1_data)
```


# Using data returned by `raustats`

Statistics returned by the `raustats` package functions are generally in long
(tidy) format data frames. This provides for easy integration with packages like
`ggplot2`, `tidyr` and `dplyr`. The following example illustrates how data
downloaded using the `raustats` can be easily transformed and plotted. This
example uses data from ABS' *Private New Capital Expenditure and Expected
Expenditure* (ABS Catalogue no. 5265.0).

First download selected tables from catalogue no. 5265.0.
```{r message=FALSE, results=FALSE}
capex_q <-
  abs_cat_stats("5625.0",
                tables=c("Actual Expenditure by Type of Asset and Industry - Current Prices",
                         "Actual Expenditure, By Type of Industry - Chain Volume Measures",
                         "Actual and Expected Capital Expenditure by Industry.+:Current Prices"))
```

Then add a new variable denoting Australian state/territory.

```{r results=FALSE, message=FALSE}
library(dplyr)
## Add state/territory variable
capex_q <- capex_q %>%
  mutate(state = sub(sprintf(".*(%s).*",
                             paste(c("New South Wales","Victoria","Queensland","South Australia",
                                     "Western Australia","Tasmania","Northern Territory",
                                     "Australian Capital Territory","Total \\(State\\)"),
                                   collapse="|")),
                     "\\1", data_item_description, ignore.case=TRUE))
```

Finally, plot quarterly time series mining sector capital expenditure (at
current prices) by Australian state and territory using `ggplot`.

```{r fig.height=4, fig.width=7, fig.path="figures/", error=TRUE, message=TRUE, verbose=TRUE}
library(ggplot2)
## Filter mining capital expenditure
capex_q_min <- capex_q %>%
  filter(grepl("mining", data_item_description, ignore.case=TRUE)) %>%
  filter(grepl("actual", data_item_description, ignore.case=TRUE)) %>%
  filter(grepl("current price", data_item_description, ignore.case=TRUE)) %>%
  filter(grepl("Total \\(Type of Asset.+\\)", data_item_description, ignore.case=TRUE))

ggplot(data=capex_q_min) +
  geom_line(aes(x=date, y=value/10^3, colour=state)) +
  scale_x_date(date_labels="%b\n%Y") +
  scale_y_continuous(limits=c(0, NA)) +
  labs(title="Australian mining sector capital expenditure, by state",
       y="Capital expenditure ($ billion)", x=NULL) +
  guides(colour = guide_legend(title=NULL)) + 
  theme(plot.title = element_text(hjust=0.5),
        legend.box = "horizontal",
        legend.position = "bottom",
        axis.text.x=element_text(angle=0, size=8))
```


# Unresolved issues

There are a few behaviours of the [ABS.Stat](http://stat.data.abs.gov.au/) API
that may help explain any unexpected results. As the
[ABS.Stat](http://stat.data.abs.gov.au/) API is still in Beta release and
subject to revision, some of these issues will be addressed in future versions
of the `raustats` package, as [ABS.Stat](http://stat.data.abs.gov.au/)
transitions towards a stable release version.


## Searching in other languages

The [ABS.Stat](http://stat.data.abs.gov.au/) API includes scope to cater to
multiple languages. However, at the time of writing, French appears to be the
only other language included in the data sets. Moreover, the text for many
records that are denoted as French appear to be in English. The ABS API calling
functions included in this library allow users to specify their preferred
language, but this functionality has been little tested to date.


## Query, data and session constraints

As already noted, [ABS.Stat](http://stat.data.abs.gov.au/) has a query string
length limit (1000 characters) and record return limit (1 million
records). Accordingly, the `abs_stats` function includes the
`enforce_api_limits` argument to catch these limits before the query is
submitted. The argument also allows the user to override the limits and submit
the query anyway. This argument may be deprecated in future versions.

The [ABS.Stat](http://stat.data.abs.gov.au/) API also has a 10-minute session
time limit for users to download datasets via the SDMX service. The functions in
`raustats` almost exclusively use the SDMX-JSON query interface, so it is not
clear whether the session time limit applies. If time limits are an issue, the ABS
advises users to submit multiple smaller requests to retrieve the required data.


# Resources

  * [ABS.Stat -- http://stat.data.abs.gov.au/](http://stat.data.abs.gov.au/)
  * [ABS.Stat FAQ](http://stat.data.abs.gov.au/)
  

<!-- -------------------------------- EOF ---------------------------------- -->
